### 2025-05-25 proposal/formalization_of_proofs.tex
Proposal drafting

\documentclass{article}
\author{ Ashwot Acharya, Bishesh Bohora, Supreme chaudary}

\title{Working Principles Of Proof Assistants And Formalization Of Some Proofs In Linear Algebra}

\begin{document}
\maketitle
\pagebreak
\section{Introduction}
Proof assistants helps with formalization of mathematical proofs in computer which enables for faster proofs computations.
Some proofs could have many many special cases which isn't easy to figureout using tradational methods, but with the use of proof assistant these cases can easily be accounted for.
Proof assistant aren't just about guiding with complicated proofs but also can help to check if the proof that is written is correct especially for those long and tedious proofs.

In 1998 Thomas C. Hales announced that he had proved the kepler conjecture in 1998 (Solane,1998) However the peer review took over 4 years especially due to the proof being incredibely difficult to check , with over a dozen of mathematician to refree the proof and although 
they were 99\% certain it was not untill the proof became formalized that they were able to completely validate the correctness of the proof, which according to Thomas Hales would have taken 20 person-year of manual work.(Szpiro, G, 2003)

\section{Problem statement}
This projects aims to design and implementat Haskell-Based proof assistant software and formalize some  of the proofs of linear Algebra. Leveraging Haskell's strong type system and its purely functional nature.
\section{Research objectives}
\begin{enumerate}
    \item To Investigate the current exsisting proof assistant software and identify the mathematics behind them.
    \item To implementat Intuitionistic type theory and apply the logic of constructivism.
    \item To formalize some selected proofs of Linear Algebra
\end{enumerate}
\section{Literature Review}

\section{Methodology}

\section{Expected Outcomes}
\section{Significiance}
\section{Work Plan}
\pagebreak
\section{Refrences}
\begin{enumerate}
    \item Sloane, N.J.A. (1998). Kepler's conjecture confirmed. Nature, 395(6701), pp.435-436 doi:https://doi.org/10.1038/26609.
    \item Szpiro, G. (2003). Does the proof stack up? Nature, 424(6944), pp.12–13,  doi:https://doi.org/10.1038/424012a.

\end{enumerate}

\end{document}

### 2025-05-30 proposal/formalization_of_proofs.tex
updated the proposal and initialized haskell file

@@
 \documentclass{article}
+
+\title{
+    {Working Principles Of Proof Assistants And Formalization Of Some Proofs in Agda}\\
+    {\large Suprevisior : K.B manandar}\\
+
+    }
     \author{ Ashwot Acharya, Bishesh Bohora, Supreme chaudary}
 
-\title{Working Principles Of Proof Assistants And Formalization Of Some Proofs In Linear Algebra}
+\begin{document}
 
-\begin{document}
 \maketitle
+
 \pagebreak
 \section{Introduction}
-Proof assistants helps with formalization of mathematical proofs in computer which enables for faster proofs computations.
-Some proofs could have many many special cases which isn't easy to figureout using tradational methods, but with the use of proof assistant these cases can easily be accounted for.
-Proof assistant aren't just about guiding with complicated proofs but also can help to check if the proof that is written is correct especially for those long and tedious proofs.
-
 In 1998 Thomas C.
@@
 they were 99\% certain it was not untill the proof became formalized that they were able to completely validate the correctness of the proof, which according to Thomas Hales would have taken 20 person-year of manual work.(Szpiro, G, 2003)
+Proof assistants helps with formalization of mathematical proofs in computer which enables for faster proofs computations.\\
+Some proofs could have many many special cases which isn't easy to figureout using tradational methods, but with the use of proof assistant these cases can easily be accounted for.
+Proof assistant aren't just about guiding with complicated proofs but also can help to check if the proof that is written is correct especially for those long and tedious proofs.
 
@@
 \section{Literature Review}
-
+Proof assistant rely extremely heavily on the type of logic it is coded on, for example 
 \section{Methodology}
+Investigating the  Coq and the agda proof assistant and assessing the Logic that each proof assistant use and how the proof assistant use such logic to help check if the proof is valid.
+We will also be inverstigating how to formalize some proofs in one of the proof assistant, and seeing how the logic holds in formalization of proofs.
 
 \section{Expected Outcomes}
+Learning indepth about type theory, curry-howard correospondance and formalization of some proofs.  
 \section{Significiance}
+Being able to Formalze proofs in compute
 \section{Work Plan}
+\begin{center}
+    \begin{tabular}{|c|c|}
+        \hline
+        Week & Work plan \\
+        \hline
+         1 & understanding the various simple type theory models \\
+        \hline
+         2 & understanding the implementation of type theory models in digital proof assistant \\
+        \hline
+         3 & Understanding Purely functional Programming paradigm and $ \lambda $-calculus\\
+        \hline
+         4 & Working Principles of agda and its core implementation \\
+        \hline 
+         5 & Implementation of some proofs in agda \\
+        \hline
+
+    \end{tabular}
+\end{center}
 \pagebreak
+
 \section{Refrences}

### 2025-05-30 proposal/formalization_of_proofs.tex
Extended Intro, problem statement and clarified objectives

@@
     {Working Principles Of Proof Assistants And Formalization Of Some Proofs in Agda}\\
-    {\large Suprevisior : K.B manandar}\\
+    {\large Supervisor : K.B manandar}\\
 
@@
 they were 99\% certain it was not untill the proof became formalized that they were able to completely validate the correctness of the proof, which according to Thomas Hales would have taken 20 person-year of manual work.(Szpiro, G, 2003)
-Proof assistants helps with formalization of mathematical proofs in computer which enables for faster proofs computations.\\
-Some proofs could have many many special cases which isn't easy to figureout using tradational methods, but with the use of proof assistant these cases can easily be accounted for.
-Proof assistant aren't just about guiding with complicated proofs but also can help to check if the proof that is written is correct especially for those long and tedious proofs.
-
+Proof assistants helps with formalization of mathematical proofs in computer which enables for their verification digitally.
+Some exhaustive proofs may contain large number of cases which is not feasible to write and verify manually and requires computer assitance.\\ 
+Such Proof assistants(systems) work by treating proofs as computational objects and checking them using strict logical rules.
+To enable this, proofs must be written in a formal language that removes ambiguity and allows machines to interpret them accurately.
+Underlying this process is a theoretical framework—often type theory—that defines how propositions and proofs relate to each other.
+\\
+This method bridges programming and logic: writing a proof resembles writing a program, and verifying it is like type-checking.
+By interpreting logic computationally, proof assistants ensure that every proof is exact, complete, and verifiable by a machine.
+This project invloves investigating this process.
+Among all available proof assistants Agda was chosen for this project due to its constructive nature and minimal yet expressive language.
 \section{Problem statement}
-This projects aims to design and implementat Haskell-Based proof assistant software and formalize some  of the proofs of linear Algebra.
-Leveraging Haskell's strong type system and its purely functional nature.
+Proof assistants are reliable tools for verification of formal mathematical theorems and computer programs.
+This project aims to analyze the theoretical foundations of such, and investigates the connection between mathematical proofs and computer programs through the Curry Howard Correspondence.
+On practice, this project will demonstrate the use of Agda, a proof assistant by formalizing some well known proofs and dissect the verification process in Agda, through lens of Dependent Type Theory on which it is based upon.
 \section{Research objectives}
 \begin{enumerate}
-    \item To Investigate the current exsisting proof assistant software and identify the mathematics behind them.
-    \item To implementat Intuitionistic type theory and apply the logic of constructivism.
-    \item To formalize some selected proofs of Linear Algebra
+    \item To explore the correspondence between Proofs and Programs, Type Theory and Intuitionstic Logic.
+    \item To investigate a current exsisting proof assistant (Agda) and identify the mathematics behind it.
+    \item To formalize some selected proofs in Agda and demonstrate verification process.
 \end{enumerate}
@@
 \section{Significiance}
-Being able to Formalze proofs in compute
+Writing proofs in a formal language introduces more rigor and reduces ambiguity.
+A formally written proof in proof assitants enables them to be treated as executable code.
+This allows it to be checked or verified mechanically.
+This way we can offload trust from a human reader to a computer which applies logic rules without oversight, reducing errors in writing and verifying proofs for mathematical theorems.
+On the other side, with the same theoretical base, treating programs written in some programming language and verifying them or checking their correctness like proofs introduces a stronger reliablity on them.
+This makes codes written for sensitive applications secure and assures that they work as intended.
+
 \section{Work Plan}
@@
         \hline
-         1 & understanding the various simple type theory models \\
+         1 & Understanding (Dependent) Type Theory and Proof Theory\\
         \hline
-         2 & understanding the implementation of type theory models in digital proof assistant \\
+         2 & Understanding the implementation of type theory models in digital proof assistant \\
         \hline
@@
         \hline
-         4 & Working Principles of agda and its core implementation \\
+         4 & Working Principles of Agda and its core implementation \\
         \hline 
-         5 & Implementation of some proofs in agda \\
+         5 & Formalization of some proofs in Agda \\
         \hline

### 2025-05-31 proposal/formalization_of_proofs.tex
rewording few lines

@@
 In 1998 Thomas C.
-Hales announced that he had proved the kepler conjecture in 1998 (Solane,1998) However the peer review took over 4 years especially due to the proof being incredibely difficult to check , with over a dozen of mathematician to refree the proof and although
-they were 99\% certain it was not untill the proof became formalized that they were able to completely validate the correctness of the proof, which according to Thomas Hales would have taken 20 person-year of manual work.(Szpiro, G, 2003)
-Proof assistants helps with formalization of mathematical proofs in computer which enables for their verification digitally.
+Hales announced that he had proved the kepler conjecture (Solane,1998) However the peer review took over 4 years especially due to the proof being incredibely difficult to check , with over a dozen of mathematician to refree the proof and although
+they were 99\% certain it was not untill the proof became formalized that they were able to completely validate the correctness of the proof, which according to Thomas Hales, would have taken 20 person-year of manual work.(Szpiro, G, 2003)
+\\Proof assistants helps with formalization of mathematical proofs in computer which enables for their verification digitally.
 Some exhaustive proofs may contain large number of cases which is not feasible to write and verify manually and requires computer assitance.\\ 

### 2025-05-31 proposal/formalization_of_proofs.tex
info about mott

--- 2025-05-30 proposal/formalization_of_proofs.tex
@@
 This project aims to analyze the theoretical foundations of such, and investigates the connection between mathematical proofs and computer programs through the Curry Howard Correspondence.
-On practice, this project will demonstrate the use of Agda, a proof assistant by formalizing some well known proofs and dissect the verification process in Agda, through lens of Dependent Type Theory on which it is based upon.
+On practice, this project will demonstrate the use of Agda, a proof assistant by formalizing some well known proofs and dissect the verification process in Agda, through lens of Martin-Löf Type Theory (MLTT) on which it is based upon.
 \section{Research objectives}

### 2025-05-31 Presentation/Presentation.tex
initated Presentation

\documentclass{beamer}

\usetheme{Berlin}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}  
\usepackage{amsmath}   
\usepackage{hyperref}  
\usepackage{url}       
\usepackage{cite}

\title{Working Principles of Proof Assitants and Formalization of Some Proofs in Agda}
\author{Ashwot Acharya, Bishesh Bohora, Supreme Chaudary}
\institute{Kathmandu University}
\date{\today}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
\end{frame}

\section{Introduction And Motivation}

\begin{frame}{Introduction}
    As the title suggests, our project will revolve around exploration of theoritical fondations behind Proof Assitants and practice them.

\end{frame}

\begin{frame}{Proof Assistants}
        In computer science and mathematical logic, a proof assistant or interactive theorem prover is a software tool to assist with the development of formal proofs by human–machine collaboration.\cite{wikipedia_proof_assistant}

        Examples:
        \begin{itemize}
            \item Coq
            \item LEAN 
            \item Agda
        \end{itemize}
\end{frame}
\begin{frame}{History}
\begin{itemize}
  \item \textbf{Gödel’s Incompleteness Theorems (1930s)}: Revealed limitations of formal systems; sparked interest in formal logic and verification.

  \item \textbf{ Computability Theory (1940s–50s)}: Turing machines and $\lambda$-calculus laid the groundwork for mechanized reasoning.
\item \textbf{Logic Theorist (1954)}: First automated theorem prover by Newell and Simon, capable of proving theorems in propositional logic.

  \item \textbf{LISP (1960)}: A symbolic programming language created by John McCarthy; became essential for early theorem proving systems.

   \item \textbf{Automath (1967)}: First system to check mathematical proofs using dependent types.

\end{itemize}
\end{frame} 
\begin{frame}

\begin{itemize}
   \item \textbf{LCF \& ML (1970s)}: Introduced tactic-based proofs and the ML programming language; foundational to later systems.

  \item \textbf{Coq (1986)}: A proof assistant based on constructive type theory, supporting verified programming and formal proofs.

  \item \textbf{Isabelle (1989)}: Generic theorem prover with support for multiple logics and strong automation tools.

  \item \textbf{Four-Color Theorem (1996)}: First major mathematical theorem re-verified by proof assistants (Coq and HOL).

  \item \textbf{Feit-Thompson Theorem (2012)}: Large-scale group theory proof formalized in Coq, showcasing proof assistant capability.

  \item \textbf{Lean (2015-2023)}: Modern proof assistant combining type theory with performance and usability; popular in formal math via \texttt{mathlib}.

\end{itemize}
\end{frame}
\begin{frame}{Motivation}
    \begin{itemize}
  \item Strong interest in mathematics and formal reasoning.
  \item Discovered type theory through internet memes on category theory.
  \item Fascinated by the Four Colour Theorem and its computer-assisted proof.
  \item The rise of AI raised the question: \textit{"How do computers understand reasoning?"}
  \item Drawn to functional programming, which closely mirrors mathematical logic and structure.

\end{itemize}

\end{frame}

\section{Glance At Theoritical Foundations}
\begin{frame}{Type Theory}
\begin{itemize}
  \item Type theory is a formal system that classifies expressions by their "types."
  \item Originally developed as an alternative to set theory for foundations of mathematics.
  \item Predecessor to Dependent Type Theory, Martin Löf Type Theory which form basis for various proof assitants.
  \item Types prevent logical paradoxes and provide a basis for constructive reasoning.
\end{itemize}
\end{frame}
\begin{frame}{Curry–Howard Correspondence}
\begin{itemize}
  \item A deep analogy between **logic and computation**:
  \begin{itemize}
    \item Propositions $\leftrightarrow$ Types
    \item Proofs $\leftrightarrow$ Programs
  \end{itemize}
  \item A proof of a proposition is a program of a corresponding type.
  \item Enables writing code that is **correct-by-construction**.
  \item Fundamental to systems like Coq, where proving a theorem is like writing a program.
\end{itemize}
\end{frame}
\begin{frame}{$\lambda$-Calculus and Functional Programming}
    \begin{itemize}
        \item \textbf{Lambda Calculus}: A minimal formal system for function definition and application; the foundation of computation theory.
  \item \textbf{Functional Programming}: Directly inspired by lambda calculus; treats computation as evaluation of mathematical functions.

  \item In proof assitants, Core logic is based on typed lambda calculus.
      \item Tools like Coq and Agda embed functional languages with type theory.
    \end{itemize}
\end{frame}

\section{Methodology and Tool}
\begin{frame}{Methodology}

\end{frame}
\begin{frame}{Agda}
Agda is a functional programming language with dependent types. It is based on Martin Löf Type Theory. And most importantly it is a proof assitant.
\cite{inproceedings}
\end{frame}
\begin{frame}{Why Agda?}
\end{frame}

\section{Work Plan}
\begin{frame}{Work Plan}
\end{frame}

\section{Significance and Expected Outcomes}
\begin{frame}{Significance}
\end{frame}
\begin{frame}{Expected Outcomes}
\end{frame}

\section{Conclusion}
\begin{frame}{References}
\bibliographystyle{apalike}       
\bibliography{references} 
\end{frame}

\end{document}

### 2025-06-01 proposal/formalization_of_proofs.tex
Literature review edited

--- 2025-05-31 proposal/formalization_of_proofs.tex
@@
 \documentclass{article}
-
+\usepackage{hyphenat}
+\usepackage{booktabs}
 \title{
     {Working Principles Of Proof Assistants And Formalization Of Some Proofs in Agda}\\
-    {\large Supervisor : K.B manandar}\\
+    {\large Supervisor: K.B manandar}\\
 
@@
 In 1998 Thomas C.
-Hales announced that he had proved the kepler conjecture (Solane,1998) However the peer review took over 4 years especially due to the proof being incredibely difficult to check , with over a dozen of mathematician to refree the proof and although
-they were 99\% certain it was not untill the proof became formalized that they were able to completely validate the correctness of the proof, which according to Thomas Hales, would have taken 20 person-year of manual work.(Szpiro, G, 2003)
+Hales announced that he had proved the kepler conjecture (Solane,1998) However the peer review took over 4 years especially due to the proof being incredibely difficult to check, with over a dozen of mathematician to refree the proof and although
+they were 99\% certain it was not untill the proof became formalized that they were able to completely validate the correctness of the proof, which according to Thomas Hales, would have taken 20 person-year of manual work.
+(Szpiro, G, 2003)
 \\Proof assistants helps with formalization of mathematical proofs in computer which enables for their verification digitally.
 Some exhaustive proofs may contain large number of cases which is not feasible to write and verify manually and requires computer assitance.\\ 
-Such Proof assistants(systems) work by treating proofs as computational objects and checking them using strict logical rules.
+Such Proof assistants (systems) work by treating proofs as computational objects and checking them using strict logical rules.
 To enable this, proofs must be written in a formal language that removes ambiguity and allows machines to interpret them accurately.
@@
 This project aims to analyze the theoretical foundations of such, and investigates the connection between mathematical proofs and computer programs through the Curry Howard Correspondence.
-On practice, this project will demonstrate the use of Agda, a proof assistant by formalizing some well known proofs and dissect the verification process in Agda, through lens of Dependent Type Theory on which it is based upon.
+On practice, this project will demonstrate the use of Agda, a proof assistant by formalizing some well known proofs and dissect the verification process in Agda,
+through lens of Dependent Type Theory on which it is based upon. 
 \section{Research objectives}
@@
 \section{Literature Review}
-Proof assistant rely extremely heavily on the type of logic it is coded on, for example 
+Proofs or sometimes called the mathematical Proofs are at the foundation of all of mathematics and no mathematical reason are accepted unless they are proven by mathematical reasoning.
+The construction of proofs begins with initiation of terms, axioms and may bed dependent on previously proven
+theorems. Each statement in mathematical proofs must be correct which means to say that each step of the proof must be logically followed from its preceeding steps. (Grami Ali, 2023)\\
+Proof assistant are comnputer system designed to do mathematics but not in the computational sense but more towards the aspect of reasoning, proving and defining.
+In proof assistant a user can setuo a mathematical theory, defines its properities and be able to perform logical task on them.
+Most proof assistant now a days are used my professionals to formalize mathematical theories and prove them which is difficult as it requires adding lot of details for it to be valid
+in the assisting software. This step of writing proofs in a formal language is called the formalization of proofs. (H Geuvers, 2009)  \\ 
+The earliest work on computer assisted proof dates back to the 1950's and 1960s 
+[Davis,1957; Gilmore, 1960; Davis and Putnam, 1960; Wang, 1960; Prawitz et al., 1960] who were devoted to automated theorem proving. 
+Automated theorem proving but Interest in more interactive arrangement began later on when full automation became a much more difficult problem than supporting human guided proofs.
+(Harrison, Urban and Wiedijk ,2014)
+Auto-Math is considered probably the earliest of the interactive theorem provers, and was the first one of the theorem Provers to use Curry-howard isomorphism.
 \section{Methodology}
@@
 \begin{center}
-    \begin{tabular}{|c|c|}
-        \hline
+    \begin{tabular}{p{1cm} p{10cm} }
+        \toprule
         Week & Work plan \\
-        \hline
+        \toprule
          1 & Understanding (Dependent) Type Theory and Proof Theory\\
-        \hline
+        \midrule
          2 & Understanding the implementation of type theory models in digital proof assistant \\
-        \hline
+        \midrule
          3 & Understanding Purely functional Programming paradigm and $ \lambda $-calculus\\
-        \hline
+        \midrule
          4 & Working Principles of Agda and its core implementation \\
-        \hline 
+        \midrule 
          5 & Formalization of some proofs in Agda \\
-        \hline
+        \bottomrule
 
@@
 \begin{enumerate}
-    \item Sloane, N.J.A. (1998). Kepler's conjecture confirmed. Nature, 395(6701), pp.435-436 doi:https://doi.org/10.1038/26609.
-    \item Szpiro, G. (2003). Does the proof stack up? Nature, 424(6944), pp.12–13,  doi:https://doi.org/10.1038/424012a.
-
+    \item Sloane, N.J.A. (1998). Kepler's conjecture confirmed. Nature, 395(6701), pp.435\hyp{}436 doi:https://doi.org/10.1038/26609.
+    \item Szpiro, G. (2003). Does the proof stack up? Nature, 424(6944), pp.12\hyp{}3,  doi:https://doi.org/10.1038/424012a
+    \item Grami, A. (2022). Proof Methods. Elsevier eBooks, [online] pp.55\hyp{}65 doi:https://doi.org/10.1016/b978-0-12-820656-0.00004-6.
+    \item Geuvers, H. (2009). Proof assistants: History, ideas and future. Sadhana, 34(1), pp.3\hyp{}25 doi:https://doi.org/10.1007/s12046-009-0001-5.
+    \item Harrison, J, Urban, J. and Wiedijk, F. (2014). History of Interactive Theorem Proving. North-Holland, pp.135\hyp{}214.\\doi:https://doi.org/10.1016/B978-0-444-51624-4.50004-6
 \end{enumerate}

### 2025-06-01 Presentation/Presentation.tex
Presentation: Problem statement, methodologies left

@@
 
-\title{Working Principles of Proof Assitants and Formalization of Some Proofs in Agda}
+\title{Working Principles of Proof Assistants and Formalization of Some Proofs in Agda}
 \author{Ashwot Acharya, Bishesh Bohora, Supreme Chaudary}
@@
 
-\section{Introduction And Motivation}
+\section{Introduction}
 
-\begin{frame}{Introduction}
-    As the title suggests, our project will revolve around exploration of theoritical fondations behind Proof Assitants and practice them.
+\begin{frame}{Problem Statement}
+    As the title suggests, our project will revolve around exploration of theoritical fondations behind Proof Assistants and practice them.
 
 \end{frame}
+\begin{frame}{Motivation}
+    \begin{itemize}
+  \item Strong interest in mathematics and formal reasoning.
+  \item Discovered type theory through internet memes on category theory.
+  \item Fascinated by the Four Colour Theorem and its computer-assisted proof.
+  \item The rise of AI raised the question: \textit{"How do computers understand reasoning?"}
+  \item Drawn to functional programming, which closely mirrors mathematical logic and structure.
 
+\end{itemize}
+
+\end{frame}
 \begin{frame}{Proof Assistants}
@@
 \end{frame}
-\begin{frame}{Motivation}
-    \begin{itemize}
-  \item Strong interest in mathematics and formal reasoning.
-  \item Discovered type theory through internet memes on category theory.
-  \item Fascinated by the Four Colour Theorem and its computer-assisted proof.
-  \item The rise of AI raised the question: \textit{"How do computers understand reasoning?"}
-  \item Drawn to functional programming, which closely mirrors mathematical logic and structure.
-
-\end{itemize}
-
-\end{frame}
 
@@
   \item Originally developed as an alternative to set theory for foundations of mathematics.
-  \item Predecessor to Dependent Type Theory, Martin Löf Type Theory which form basis for various proof assitants.
+  \item Predecessor to Dependent Type Theory, Martin Löf Type Theory which form basis for various proof assistants.
   \item Types prevent logical paradoxes and provide a basis for constructive reasoning.
@@
 
-  \item In proof assitants, Core logic is based on typed lambda calculus.
+  \item In proof assistants, Core logic is based on typed lambda calculus.
       \item Tools like Coq and Agda embed functional languages with type theory.
@@
 \begin{frame}{Agda}
-Agda is a functional programming language with dependent types. It is based on Martin Löf Type Theory. And most importantly it is a proof assitant.
+Agda is a functional programming language with dependent types. It is based on Martin Löf Type Theory. And most importantly it is a proof assistants.
 \cite{inproceedings}
@@
 \begin{frame}{Work Plan}
+  \begin{center}
+    \begin{tabular}{|c|c|}
+        \hline
+        Week & Work plan \\
+        \hline
+         1 & Understanding (Dependent) Type Theory and Proof Theory\\
+        \hline
+         2 & Understanding the implementation of type theory models in digital proof assistant \\
+        \hline
+         3 & Understanding Purely functional Programming paradigm and $ \lambda $-calculus\\
+        \hline
+         4 & Working Principles of Agda and its core implementation \\
+        \hline 
+         5 & Formalization of some proofs in Agda \\
+        \hline
+
+    \end{tabular}
+\end{center}
 \end{frame}
@@
 \begin{frame}{Significance}
+  \begin{itemize}
+    \item Writing proofs in a formal language introduces more rigor and redcues ambiguity, and makes it executable allowing computer verification.
+    \item Simplifies verification process and makes it error free.
+     \item Formalizing some proofs can be difficult, this creates a challange which gives rise to novel ideas.
+    \item Opportunity to delve into Constructivism.
+    \item Can use same theory to check correctness of computer programs/software which helps in writing secure code. 
+  \end{itemize}
+
 \end{frame}
 \begin{frame}{Expected Outcomes}
+  \begin{itemize}
+    \item Improved understanding of proofs, logic and programming.
+    \item Experience of Using A Proof Assistant and incorporating them in regular studies.
+    \item New Formalization of Proofs of various fields in Agda 
+    \item Proofs for Various Algorithms
+    \item (Re)Discovery of limitations of such proof assistants and formal language.
+    \item Future prospects and discussion on possibility of integration of formal language in Machine Learning,
+  \end{itemize}
 \end{frame}
 
-\section{Conclusion}
+\section{References}
 \begin{frame}{References}

### 2025-06-01 proposal/formalization_of_proofs.tex
Literature curry-howard correospondance

@@
 \usepackage{booktabs}
+\usepackage{a4}
 \title{
@@
 \section{Literature Review}
-Proofs or sometimes called the mathematical Proofs are at the foundation of all of mathematics and no mathematical reason are accepted unless they are proven by mathematical reasoning.
+\indent{}
+Proofs or sometimes called the mathematical Proofs are at the foundation of all 
+of mathematics and no mathematical reason are accepted unless they are proven by mathematical reasoning.
 The construction of proofs begins with initiation of terms, axioms and may bed dependent on previously proven
@@
 [Davis,1957; Gilmore, 1960; Davis and Putnam, 1960; Wang, 1960; Prawitz et al., 1960] who were devoted to automated theorem proving. 
-Automated theorem proving but Interest in more interactive arrangement began later on when full automation became a much more difficult problem than supporting human guided proofs.
-(Harrison, Urban and Wiedijk ,2014)
-Auto-Math is considered probably the earliest of the interactive theorem provers, and was the first one of the theorem Provers to use Curry-howard isomorphism.
+ Interest in more interactive arrangement began much later on when full automated system became a very difficult problem than supporting human guided proofs.\\
+AutoMath is considered possibly the earliest of human guided theorem provers which gave rise to various other systems of theorem prover and is the first program to use curry howard isomorphism for encoding of proofs (Harrison, Urban and Wiedijk ,2014) \\
+Curry-Howard isomorphism states that any dertivation written in Intuitionstic Logic can be writing in their correosponding typable $ \lambda$\hyp{}terms.
+which means that proofs can be represented in programing terms, In short an isomorphism can be made between programs and logical proofs (Sørensen, M.H.
+and Urzyczyin, 2006).
+Purely functional programming languages are based on $\lambda$-calculus (H.P Barendregt, 1990), this enables computability of forementioned derivations or proofs.
+
 \section{Methodology}
-Investigating the  Coq and the agda proof assistant and assessing the Logic that each proof assistant use and how the proof assistant use such logic to help check if the proof is valid.
+
+In this project we will be investigating the  Coq and the agda proof assistant and assessing the Logic that each proof assistant use and how \the proof assistant use such logic to help check if the proof is valid.
 We will also be inverstigating how to formalize some proofs in one of the proof assistant, and seeing how the logic holds in formalization of proofs.
@@
     \item Szpiro, G. (2003). Does the proof stack up? Nature, 424(6944), pp.12\hyp{}3,  doi:https://doi.org/10.1038/424012a
-    \item Grami, A. (2022). Proof Methods. Elsevier eBooks, [online] pp.55\hyp{}65 doi:https://doi.org/10.1016/b978-0-12-820656-0.00004-6.
+    \item Grami, A. (2022). Proof Methods. Elsevier eBooks, [online] pp.55\hyp{}65\\doi:https://doi.org/10.1016/b978-0-12-820656-0.00004-6.
     \item Geuvers, H. (2009). Proof assistants: History, ideas and future. Sadhana, 34(1), pp.3\hyp{}25 doi:https://doi.org/10.1007/s12046-009-0001-5.
     \item Harrison, J, Urban, J. and Wiedijk, F. (2014). History of Interactive Theorem Proving. North-Holland, pp.135\hyp{}214.\\doi:https://doi.org/10.1016/B978-0-444-51624-4.50004-6
+    \item Sørensen, M.H.
+and Urzyczyin, P.
+(2006).
+The Curry-Howard isomorphism.
+Studies in Logic and the Foundations of Mathematics, pp.77\hyp{}101.\\doi:https://doi.org/10.1016/s0049-237x(06)80005-4.
+    \item BARENDREGT, H.P. (1990). Functional Programming and Lambda Calculus. Formal Models and Semantics, pp.321\hyp{}363 \\doi:https://doi.org/10.1016/b978-0-444-88074-1.50012-3.
+
 \end{enumerate}

### 2025-06-03 Presentation/Presentation.tex
Finishing up the project proposal presentation small edits in proofs

--- 2025-05-31 Presentation/Presentation.tex
@@
 \usepackage{cite}
+\usepackage{booktabs}
 
@@
 \begin{frame}{Methodology}
-
+    \begin{itemize}
+      \item Investigating the Coq and the agda proof assistant
+      \item Assessing the logic behind each of these proof assistants
+      \item Investigating How to formalized proofs in Agda
+    \end{itemize}
 \end{frame}
@@
 \begin{frame}{Why Agda?}
+  \begin{itemize}
+    \item Dependently Typed programming language
+    \item Fully embraces the Curry–Howard isomorphism
+    \item Minimal where users typically define the core logical construct from small set of primitives
+    \item Active Community
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Work Plan}
+      \begin{tabular}{p{1cm} p{8cm} }
+        \toprule
+        Week & Work plan \\
+        \toprule
+         1 & Understanding (Dependent) Type Theory and Proof Theory\\
+        \midrule
+         2 & Understanding the implementation of type theory models in digital proof assistant \\
+        \midrule
+         3 & Understanding Purely functional Programming paradigm and $ \lambda $-calculus\\
+        \midrule
+         4 & Working Principles of Agda and its core implementation \\
+        \midrule 
+         5 & Formalization of some proofs in Agda \\
+        \bottomrule
+
+    \end{tabular}
 \end{frame}
@@
 \begin{frame}{Significance}
+  \begin{itemize}
+    \item Writing proofs in a formal language adds rigor and reduces ambiguity, allowing proof assistants to treat proofs as executable code that can be mechanically checked, shifting trust from human readers to computers and minimizing errors in mathematical verification.
+
+   \item Using the same theoretical foundation, treating programs as proofs and verifying their correctness increases reliability, ensuring that code—especially for sensitive applications—is secure and functions exactly as intended.
+
+  \end{itemize}
 \end{frame}
 \begin{frame}{Expected Outcomes}
+  \begin{itemize}
+    \item Understanding of dependent type theory and logics
+    \item Understanding of $\lambda$-calculus and purely functional programming
+    \item  Formalization of some selected proofs in agda
+  \end{itemize}
 \end{frame}

### 2025-07-03 Mainpaper/paper.tex
initializing main paper

\documentclass[12pt]{article}

% ======== Packages =========
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{geometry}
\geometry{a4paper, margin=1in}
\usepackage{titlesec}
\titleformat{\section}{\large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{1em}{}

\title{Working Principles of Proof Assistants and Formalization of some proofs in Agda}
\author{Bishesh Bohora, Supreme Chaudhary, Ashwot Acharya \\
\small Kathmandu University 
\small \texttt{}
}
%mention supervisor
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
\end{abstract}

\tableofcontents
\newpage

\section{Introduction}

\section{Foundations}

\subsection{Logic Foundations}
This works assumes prior knowledge of Propositional and Predicate Logic. 
\subsubsection{Natural Deduction}
The propositions or formulas in Propositional Logic can be verified or proved simply by constructing their truth tables. But for logically complex propositions or propositions with many atomic statements, it becomes difficult to construct a truth table. With predicates, this becomes impossible. Therefore, to mitigate this we adhere to a basic set of inference rules with which we derive conclusions from assumptions in step by step, structured manner. The rule based system which allows us to reason about logical structure of propositions is known as \textbf{Natural Deduction}. \\
With the rules in (refer sth, either appendix or a book) we now present an example on how a proof is carried out,
(write a proof here).(also use law of excluded middle here)

The above works for propositional case, now adding these  Introduction and Elimination rules for quantifiers, we get a set of rules enough for predicate logic.
(write extra rules) as in (refer)

and a proof as (proof with quantifiers)

The \textbf{soundness} and \textbf{completeness} of this system are discussed here (refer). 
\subsubsection{Intuitionstic Logic}

\section{Main Content}
Divide into multiple subsections as appropriate. For example:
\subsection{Topic A}
Discuss and explain this sub-topic clearly.

\subsection{Topic B}
Likewise, expand on another subtopic. Include figures, equations, or examples to aid exposition.

\section{Comparative Discussion}
(Optional) If your report compares different approaches, schools of thought, or perspectives, lay that out here.

\section{Applications or Case Studies}
(Optional) Describe practical uses or illustrative examples, if any.

\section{Summary and Conclusion}
Summarize the key points discussed. Mention what the reader should take away, and possibly directions for further reading or study.

% ======== References =========
\bibliographystyle{plain} % Or your preferred style
\bibliography{references}

\end{document}

### 2025-07-03 Mainpaper/paper.tex
basic outline

@@
 
-\section{Main Content}
-Divide into multiple subsections as appropriate. For example:
-\subsection{Topic A}
-Discuss and explain this sub-topic clearly.
+\section{General Architecture of Proof Assistants}
 
-\subsection{Topic B}
-Likewise, expand on another subtopic. Include figures, equations, or examples to aid exposition.
+\section{Upon Some Proof Assistants and Comparative Study}
+\subsection{Coq}
 
-\section{Comparative Discussion}
-(Optional) If your report compares different approaches, schools of thought, or perspectives, lay that out here.
+\subsection{Lean}
 
-\section{Applications or Case Studies}
-(Optional) Describe practical uses or illustrative examples, if any.
+\subsection{Isabelle}
 
-\section{Summary and Conclusion}
-Summarize the key points discussed. Mention what the reader should take away, and possibly directions for further reading or study.
+\subsection{Agda}
+
+\section{ Agda}
+
+\section{Formalization of Some Proofs}
 
 % ======== References =========
-\bibliographystyle{plain} % Or your preferred style
+\bibliographystyle{plain} 
 \bibliography{references}

### 2025-07-04 Mainpaper/paper.tex

@@
 \usepackage{geometry}
+\usepackage{bussproofs}
+
 \geometry{a4paper, margin=1in}
@@
 \section{Foundations}
-
+This section sheds light upon some fundamental concept on which Proof Assistants work along with how a proof should be written so that it can be mechanically verified.
 \subsection{Logic Foundations}
-This works assumes prior knowledge of Propositional and Predicate Logic. 
+This works assumes prior knowledge of Propositional and Predicate Logic. And we refer to both together as \textbf{Classical}.
 \subsubsection{Natural Deduction}
@@
 \\
-With the rules in (refer sth, either appendix or a book) we now present an example on how a proof is carried out,
-(write a proof here).(also use law of excluded middle here)
+With the rules in Section 2.3 of \cite{Alrubyli2021}  we now present examples on how a proof is carried out with Natural Deduction for Classical Logic.
 
-The above works for propositional case, now adding these  Introduction and Elimination rules for quantifiers, we get a set of rules enough for predicate logic.
-(write extra rules) as in (refer)
+\newtheorem{example}{Example}[section] 
+\begin{example}
+    \label{contra}
+    \textbf{$(A \land\neg A) \to \bot $ }
 
-and a proof as (proof with quantifiers)
+    By negation introduction we can write $\neg A$ as $A \to \bot$
 
-The \textbf{soundness} and \textbf{completeness} of this system are discussed here (refer). 
+    $(A \land (A \to \bot)) \to \bot$
+
+    \begin{prooftree}
+        \AxiomC{$[A]$}
+        \AxiomC{$[A \to \bot]$}
+        \RightLabel{\scriptsize$\to$E}
+        \BinaryInfC{$\bot$}
+    \end{prooftree}
+
+\end{example}
+
+\begin{example}
+    \label{lem}
+    Proof for Law of Excluded Middle ($P \lor \neg P$)\\
+    Note that Example \ref{contra}'s result is used here.
+    \begin{prooftree}
+\AxiomC{$[P]$}
+\RightLabel{\scriptsize$\lor$I$$}
+\UnaryInfC{$P \lor \neg P$}
+\AxiomC{$[\neg (P \lor \neg P)]$}
+\RightLabel{\scriptsize$\bot$}
+\BinaryInfC{$\bot$}
+\RightLabel{\scriptsize$\neg I$}
+\UnaryInfC{$\neg P$}
+\RightLabel{\scriptsize$\lor$I$$}
+\UnaryInfC{$P \lor \neg P$}
+\AxiomC{$[\neg (P \lor \neg P)]$}
+\RightLabel{\scriptsize$\bot$}
+\BinaryInfC{$\bot$}
+\RightLabel{\scriptsize$\neg$I$$}
+\UnaryInfC{$\neg \neg (P \lor \neg P)$}
+\RightLabel{\scriptsize$\neg\neg$E}
+\UnaryInfC{$P \lor \neg P$}
+\end{prooftree}
+\end{example}
+
+\begin{example}
+    $$ \forall x (A(x) \to B(x) ) , A(s)  \vdash \exists  x B(x) $$
+
+    \begin{prooftree}
+        \AxiomC{$[ \forall x (A(x) \to B(x))]$}
+        \AxiomC{$[A(s)]$}
+        \RightLabel{\scriptsize $\forall E$}
+        \BinaryInfC{$A(s) \to B(s)$}
+        \RightLabel{\scriptsize $\to E$}
+        \UnaryInfC{B(s)}
+        \RightLabel{\scriptsize $\exists I$}
+        \UnaryInfC{$\exists x B(x)$}
+    \end{prooftree}
+
+\end{example}
+
+The \textbf{soundness} and \textbf{completeness} of this system are discussed in Section 3.1 and 3.2 \cite{Alrubyli2021}. 
 \subsubsection{Intuitionstic Logic}
+Intuitionstic Logic was introduced to formalize the constructive method to do mathematics.
+Unlike in Classical Logic a statement is True if we can construct a proof for it and to claim a statement is False, again a proof of its falsity is required.
+This allows the case that some statements are not provable.
+To show something exists one must provide an method or algorithm to construct it. Proof Assistants leverage this fact.
+The constructive view of doing mathematics gives a stricter criteria. Intuitionstic  Logic can be obtained by restricting certain parts of Classical Logic, like the Law of Excluded Middle.
+
+For inference rules for Natural Deduction in Intuitionstic Logic See 2.1 \cite{Pfenning2004}, here the language is slightly different from above, we have
+\\
+\textit{Terms} \quad $t ::= x \mid a \mid f(t_1, \ldots, t_n)$
+\\
+\textit{Propositions} \quad
+$A ::= P(t_1, \ldots, t_n) \mid A_1 \land A_2 \mid A_1 \supset A_2 \mid A_1 \lor A_2 \mid \neg A \mid \bot \mid \top \mid \forall x. A \mid \exists x. A$
+
+ The main focus is that this new set of rules does not contain the double negation rule which is present in what we introduced in Section 2.1.1.
+\textbf {Example \ref{lem}} uses the Double negation rule in its proof, which we don't have now.
+This agrees to that Law of Excluded Middle does not work in Intuitionstic Logic.
+The method of proof by contradiction also relies on this rule, so constructivists omit it.\\
+These rules are revised with localized hypotheses i.e we use the above rules under a set of premises, and refer the derivation as "derived under a context". With introduction of contexts 
+$$Contexts  \ \Gamma ::= .|\Gamma,u: A$$
+[See 2.3 \cite{Pfenning2004}]
+
+\subsection{Type Theory}
 
@@
 
+\section{Challenges and Workarounds}
+
 % ======== References =========
-\bibliographystyle{plain} 
+\bibliographystyle{apalike} 
 \bibliography{references}

### 2025-07-09 test_assistant/app/MAlonzo/Code/Qhello.hs
started some basic proof writing

@@ -1,0 +1,54 @@
+{-# LANGUAGE BangPatterns #-}
+{-# LANGUAGE EmptyCase #-}
+{-# LANGUAGE EmptyDataDecls #-}
+{-# LANGUAGE ExistentialQuantification #-}
+{-# LANGUAGE NoMonomorphismRestriction #-}
+{-# LANGUAGE OverloadedStrings #-}
+{-# LANGUAGE PatternSynonyms #-}
+{-# LANGUAGE RankNTypes #-}
+{-# LANGUAGE ScopedTypeVariables #-}
+
+{-# OPTIONS_GHC -Wno-overlapping-patterns #-}
+
+module MAlonzo.Code.Qhello where
+
+import MAlonzo.RTE (coe, erased, AgdaAny, addInt, subInt, mulInt,
+                    quotInt, remInt, geqInt, ltInt, eqInt, add64, sub64, mul64, quot64,
+                    rem64, lt64, eq64, word64FromNat, word64ToNat)
+import qualified MAlonzo.RTE
+import qualified Data.Text
+
+-- hello.an-implication
+d_an'45'implication_8 ::
+  () ->
+  () ->
+  () ->
+  (AgdaAny -> AgdaAny) ->
+  (AgdaAny -> AgdaAny) -> (AgdaAny -> AgdaAny) -> AgdaAny -> AgdaAny
+d_an'45'implication_8 ~v0 ~v1 ~v2 ~v3 v4 v5 v6
+  = du_an'45'implication_8 v4 v5 v6
+du_an'45'implication_8 ::
+  (AgdaAny -> AgdaAny) -> (AgdaAny -> AgdaAny) -> AgdaAny -> AgdaAny
+du_an'45'implication_8 v0 v1 v2 = coe v1 (coe v0 v2)
+-- hello.another-implication
+d_another'45'implication_24 ::
+  () ->
+  () ->
+  () ->
+  (AgdaAny -> AgdaAny) ->
+  (AgdaAny -> AgdaAny) -> (AgdaAny -> AgdaAny) -> AgdaAny -> AgdaAny
+d_another'45'implication_24 ~v0 ~v1 ~v2 v3 ~v4 v5 v6
+  = du_another'45'implication_24 v3 v5 v6
+du_another'45'implication_24 ::
+  (AgdaAny -> AgdaAny) -> (AgdaAny -> AgdaAny) -> AgdaAny -> AgdaAny
[…]

### 2025-07-09 Mainpaper/paper.tex
till simpleTT

@@
 \usepackage{bussproofs}
-
+\usepackage{tasks}
 \geometry{a4paper, margin=1in}
@@
 
-\subsection{Type Theory}
+\subsection{$\lambda$ - Calculus and Type Theory}
+
+\subsubsection{Untyped $\lambda$-Calculus}
+It is a model of computation introduced by Alzano Church. It consists of construction and operations on lambda terms. 
+Among the lambda terms we have,
+\begin{enumerate}
+    \item Variables : $x,y,z ...$ are lambda terms
+    \item Application : If E,F are lambda terms EF is a lambda term 
+    \item Abstraction : If E is a lambda term $\lambda x. E $ is a lambda term
+\end{enumerate}
+Therefore lambda calculus is a formal system involving Abstraction and Application of functions along with some reduction rules.
+Abstraction is the definition of a anonymous function , for example \\
+$$ \lambda x.x^2 $$
+
+Now Application is calling that given function by applying it.
+
+\begin{align*}
+ (\lambda x.x^2) 3\\
+ 3^2 \\
+ 9
+\end{align*}
+The reduction rules are :
+
+\begin{enumerate}
+    \item $\alpha$-Conversion \\
+    \indent The function remains the same if all bound variable is renamed.
+    In above example $\lambda y. y^2$ does the same thing.
+
+$\lambda x.t$ is same as $ \lambda y.t[x:=y]$
+    \item $\beta$-Reduction\\
+     This is substitution. During application all the instances of bound variable in the expression is replaced by the argument. Again referring above example $x^2$ becomes $3^2$.\\
+     $(\lambda x.E) F$ is $E[x := F]$, where $E[x := F]$
+\end{enumerate}
+For comprehensive treatment of this topic, see \cite{Rojas2015}. And for the original formulation by Church \cite{}
+This is not sufficient for the mentioned application.
+
+\subsubsection{Type Theory}
+Type Theory is a formal system that associates every object or \textbf{term} a \textbf{Type}.
+The earliest version of Type Theory was introduced by Russell to avoid paradoxes in Set Theory (like \textbf{Russell's Paradox}).
+Later with development of Simply Typed Lambda Calculus, this became important as a foundation of Programming Languages.\\
+
+Unlike Sets, Types do not talk about the semantics but rather the structure or syntax only.
+For example, in Type theory, $1+2:\mathbb{N}$ we carry out the judgement that 1+2 is a Natural Number whereas while talking bout sets $1+2 \in \mathbb{N}$  means that 1+2 is inside $\mathbb{N}$.
+\\With types, we do not care what 1+2 means, the information we have is the operator '+' has a type $\mathbb{N} \to \mathbb{N}$ and  $1:\mathbb{N}, 2:\mathbb{N}$ therefore the judgement  $1+2:\mathbb{N}$ follows.
+Whereas with sets, we proceed as $1+2=3\in \mathbb{N}$.\\
+
+Note that verifying if a term a is of type A is decidable, i.e there is an algorithm for it.
+This process is called Typing or Type Checking.
+Proof Assistants leverage this for proof checking.
+Proof checking is decidable; proof finding not.
+\cite{Geuvers2009}\\
+
+In simple type theory there are, 
+\begin{itemize}
+    \item \textbf{Base Types} : Primitive types which are not built from anything else, Eg: Boolean Type, Nat Type etc.
+These are generally denoted by capital letters M,N,T etc.
+and we can carry out the judgement as\\
+        a:A, a is term of Type A.\\
+
+    \item \textbf{Arrow Types} : Also called as function type which take input of a type and return another.
+As we have discussed '+' has type $\mathbb{N} \to \mathbb{N}$.
+More generally if a function f takes input of type A and returns output of type B then, $f:A \to B$.
+In logic these are equivalent to implication.
+We will discuss about this correspondence later.
+
+\end{itemize}
+
+This is not expressive enough for doing constructive mathematics. So the system is extended as ,
+
+\begin{itemize}
+    \item \textbf{Product Types} 
+\end{itemize}
+
+Simply Typed lambda Calculus
+
+Dependent Type Theory
+
+\subsection{Curry Howard Correspondence}
+
+% Refer
+
+% Make a Table
+
+% Show a model of a proof 
+% one basic one mathy
+
+ \subsection{Martin Lof Type Theory }
+
+% Finally rigidify the basics refer to the original paper
+
+% Update based on knowledge.....
 
@@
 \bibliographystyle{apalike} 
-\bibliography{references}
+\bibliography{references.bib}
 

### 2025-07-09 Mainpaper/paper.tex
till simpleTLC

@@
 \usepackage{bussproofs}
-\usepackage{tasks}
+\usepackage{braket}
 \geometry{a4paper, margin=1in}
@@
 [See 2.3 \cite{Pfenning2004}]
+A small example using this set of rules,
+$$ $$
+\begin{example}
+
+\end{example}
 
@@
 
-\subsubsection{Untyped $\lambda$-Calculus}
+\subsubsection{ $\lambda$-Calculus}
 It is a model of computation introduced by Alzano Church. It consists of construction and operations on lambda terms. 
@@
 
+\begin{example}
 \begin{align*}
@@
 \end{align*}
+\end{example}
+The functions can be constructed such that they can take multiple values too, say a function to compute sum of squares can be implemented with lambda calculus as
+
+\begin{example}
+    $$\lambda x \lambda y. x^2 +y^2$$
+Its application is as 
+$$(\lambda x \lambda y. x^2 +y^2) (3) (4)$$
+$$(\lambda y. 3^2 + y^2)(4)$$
+$$3^2+4^2$$
+$$9+16$$
+$$25$$
+\end{example}
+
+This is called \textbf{currying}.\\
 The reduction rules are :
@@
 For comprehensive treatment of this topic, see \cite{Rojas2015}. And for the original formulation by Church \cite{}
-This is not sufficient for the mentioned application.
 
@@
 For example, in Type theory, $1+2:\mathbb{N}$ we carry out the judgement that 1+2 is a Natural Number whereas while talking bout sets $1+2 \in \mathbb{N}$  means that 1+2 is inside $\mathbb{N}$.
-\\With types, we do not care what 1+2 means, the information we have is the operator '+' has a type $\mathbb{N} \to \mathbb{N}$ and  $1:\mathbb{N}, 2:\mathbb{N}$ therefore the judgement  $1+2:\mathbb{N}$ follows.
+\\With types, we do not care what 1+2 means, the information we have is the operator '+' has a type $\mathbb{N} \times \mathbb{N}  \to \mathbb{N}$ and  $1:\mathbb{N}, 2:\mathbb{N}$ therefore the judgement  $1+2:\mathbb{N}$ follows.
 Whereas with sets, we proceed as $1+2=3\in \mathbb{N}$.\\
@@
 \begin{itemize}
-    \item \textbf{Base Types} : Primitive types which are not built from anything else, Eg: Boolean Type, Nat Type etc.
+    \item \textbf{Base Types} : Primitive types which are not built from anything else, Eg: Boolean Type, Nat Type, $\bot$ or Empty Type etc.
 These are generally denoted by capital letters M,N,T etc.
@@
 
-This is not expressive enough for doing constructive mathematics. So the system is extended as ,
+Which can be further extended as,
+\begin{itemize}
+    \item \textbf{Product Types} Product types are similar to Cartesian Products from Set Theory, the terms inhabiting the product type is a pair of terms from the involving types, $a:A, b:B $ then $ \left\langle a,b \right\rangle : A \times B$ .
+    \item \textbf{Sum Type} This is similar to disjoint union with sets, therefore the term has one of the participating types. 
+     $$a:A, b:B, a:A+B, b:A+B$$
 
+\end{itemize}
+\subsubsection{Simply Typed Lambda Calculus}
+ We have already discussed about the Lambda Calculus as well as Types, now we assign a type to each lambda term.
+The functions in Untyped lambda Calculus are too general.
+In Example 2.5 we did not mention what squaring means, can we apply it to some arbitary type of data, what will the function return, these concerns remain unanswered.
+But when we assign every term a type, it can be assigned what can the function take and return.
  \begin{itemize}
-    \item \textbf{Product Types} 
+    \item Variables  x:T , a variable which has a type T
+    \item Abstraction $\lambda x:M.t:T$ , the function takes an input of type M and returns t of type T.
+    \item Application $(\lambda x:M.t:T)(a:M)$
+
  \end{itemize}
 
-Simply Typed lambda Calculus
+ We present a typed version of Example 2.5\\
+ \begin{example}
+    $$\lambda x:\mathbb{N}.x^2:\mathbb{N}$$
+    Note that now this function can only take natural numbers,\\
+    Application
+    $$(\lambda x:\mathbb{N}.x^2:\mathbb{N})(3)$$
+    which gives, along with the judgement
+    $$3^2:\mathbb{N}$$
+ \end{example}
 
-Dependent Type Theory
+\subsection{Proof Terms and The Curry Howard Correspondence}
 
-\subsection{Curry Howard Correspondence}
-
-% Refer
-
-% Make a Table
-
-% Show a model of a proof 
-% one basic one mathy
+\subsection{Dependent Type Theory and Dependently Typed $\lambda$-Calculus}
+In this extension of theory of types, we allow types to depend upon some value. For example, modeling vectors with types, we require it to depend on Natural number for its dimension. 
+u:Vec(n) means u is a vector type with n dimension and n:$\mathbb{N}$. 
+This can be taken as a function type with $Vec:\mathbb{N} \to Type$ , which takes a type of natural number as input and returns a type.
 

### 2025-07-25 Mainpaper/paper.tex
curryhow

@@
 \usepackage{titlesec}
+\usepackage{tikz}
+\usetikzlibrary{arrows.meta, positioning}
 \titleformat{\section}{\large\bfseries}{\thesection}{1em}{}
@@
 \\
-With the rules in Section 2.3 of \cite{Alrubyli2021}  we now present examples on how a proof is carried out with Natural Deduction for Classical Logic.
+With the rules in Section 2.3 of \cite{Alrubyli2021}  we now present proofs of some propositions carried out with Natural Deduction for Classical Logic.
 
+\newtheorem{proposition}{Proposition}[section] 
 \newtheorem{example}{Example}[section]
-\begin{example}
+\begin{proposition}
     \label{contra}
@@
 
-\end{example}
+\end{proposition}
 
-\begin{example}
+\begin{proposition}
     \label{lem}
     Proof for Law of Excluded Middle ($P \lor \neg P$)\\
-    Note that Example \ref{contra}'s result is used here.
+    Note that Proposition \ref{contra} is used here.
     \begin{prooftree}
@@
 \end{prooftree}
-\end{example}
+\end{proposition}
 
-\begin{example}
+\begin{proposition}
     $$ \forall x (A(x) \to B(x) ) , A(s)  \vdash \exists  x B(x) $$
@@
 
-\end{example}
+\end{proposition}
 
@@
 [See 2.3 \cite{Pfenning2004}]
-A small example using this set of rules,
-$$ $$
-\begin{example}
-
-\end{example}
 
@@
 
-\subsection{Proof Terms and The Curry Howard Correspondence}
-
-\subsection{Dependent Type Theory and Dependently Typed $\lambda$-Calculus}
+\subsubsection{Dependent Type Theory and Dependently Typed $\lambda$-Calculus}
 In this extension of theory of types, we allow types to depend upon some value. For example, modeling vectors with types, we require it to depend on Natural number for its dimension. 
@@
 This can be taken as a function type with $Vec:\mathbb{N} \to Type$ , which takes a type of natural number as input and returns a type.
+This allows indexing of types and is useful for representing predicates.
+And similarly as with simply typed $\lambda$-Calculus we can have lambda terms with dependent types.
+
+\subsection{Proof Terms and The Curry Howard Correspondence}
+    Let us treat every proposition or Logical Formula as a Type, and their proofs as the term inhabiting it.
+Intuitionstically, we interpret a proposition being true if we present a proof for it.
+Here, it will be true if there is a term (proof) inhabiting the type (proposition).
+    p:P is now interpreted as p is a proof of P. In \cite{How} the strong correspondence between intuitionstic derivation and lambda terms was mentioned. In essence the corrrespondence is as follows,
+    \begin{table}[h]
+\centering
+\begin{tabular}{|l|l|l|}
+\hline
+\textbf{Logic}                & \textbf{Computation}         & \textbf{Type Theory}             \\ \hline
+Proposition                  & Type                         & Type                             \\ \hline
+Proof                        & Program                      & Term                             \\ \hline
+Proof Normalization          & Program Evaluation           & Reduction                        \\ \hline
+Implication ($A \rightarrow B$) & Function Type ($A \to B$)    & $\lambda$-abstraction            \\ \hline
+Conjunction ($A \land B$)    & Product Type ($A \times B$)  & Pairing                          \\ \hline
+Disjunction ($A \lor B$)     & Sum Type ($A + B$)           & Tagged union                     \\ \hline
+Falsehood ($\bot$)           & Empty Type                   & No term                          \\ \hline
+Truth ($\top$)               & Unit Type                    & Singleton term                   \\ \hline
+Negation ($\neg A$)          & $A \to \bot$                 & Function to empty type           \\ \hline
+Universal Quantifier ($\forall x. A(x)$) & Dependent Product ($\Pi x : A. B(x)$) & Function over types \\ \hline
+Existential Quantifier ($\exists x. A(x)$) & Dependent Sum ($\Sigma x : A. B(x)$) & Pair with dependent type \\ \hline
+
+\end{tabular}
+\caption{Curry–Howard Correspondence: Logic, Computation, and Type Theory}
+\end{table}\\
+The use of dependent product and sum for the Quantifier highlights the necessity of Dependent Type Theory.
+The rules  (Proof Terms only), equivalent to Deduction rules of Intuitionstic Logic are in Section 2.4 of \cite{Pfenning2004}.
+Note that this does not contain the rule for predicates.
+While writing a proof in a proof assistant our reasoning follows deduction rules of this kind.
+Then to write a  program (equivalent to proof) of the proposition $A \to B \to A$, we need to construct a function that takes term of type A , term of type B and returns term of type A, that is $\lambda a{:}A.\lambda b{:}B.t{:}A $ which proceeds as
+
+   \begin{prooftree}
+  \AxiomC{}
+  \RightLabel{\scriptsize var}
+  \UnaryInfC{$a{:}A,\ b{:}B \vdash a : A$}
+  \RightLabel{\scriptsize lam}
+  \UnaryInfC{$a{:}A \vdash \lambda b{:}B.\ a : B \rightarrow A$}
+  \RightLabel{\scriptsize lam}
+  \UnaryInfC{$\vdash \lambda a{:}A.\lambda b{:}B.\ a : A \rightarrow B \rightarrow A$}
+\end{prooftree}
 
  \subsection{Martin Lof Type Theory }
+
+ \subsection{Evolution of Type Theories}
 

### 2025-07-26 Mainpaper/paper.tex
foundations

@@
 
- \subsection{Martin Lof Type Theory }
+ \subsection{Martin-Löf Type Theory }
+    Martin-Löf Type Theory is the formal logic system that was developed to support constructive mathematics.
+It's often referred to as Intuitionstic Type Theory.
+It carries the use of dependent types and is the backbone of Proof Assistants.\\
+        The following types are also valid in this system.
+        \begin{itemize}
+            \item $\prod$-Type : The dependent function type, as mentioned in Table 1. it supports the interpretation of Universal Quantifier.
+            \item $\sum$-Type : The dependent sum type, equivalent to the Existential Quantifier.
+            \item Identity Type : Equality itself is internalized as a type in this system.
+Hence, we can judge if two terms are the same.
+By the correspondence, proofs are terms then this allows proof assistants to match proofs and declare their Equality.
 
- \subsection{Evolution of Type Theories}
+        \end{itemize}
+        Furthermore, We also have \textbf{Type Universes}, Every Type itself can be assigned a Type.
+This creates a hierarchy of types.
+For example, $1:\mathbb{N},\mathbb{N}:Type:...$, The number 1 has type Natural, then Natural has a type called Type, which itself can have another type.
+Not to be interpreted as , a set being contained in another set, this gives rise to paradoxes.
+But rather take it as hierarchy of Universes.
+From this we can talk about Types themselves within our system, and  allow proof assistants to safely handle abstraction over types.\\
+        The results from  \cite{martinlof1972intuitionistic_theory} where this system was first introduced by Martin Löf, will be mentioned as necessary.
 

### 2025-07-29 Mainpaper/archi.tex

% proof_assistants_architecture.tex
\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb,url,hyperref,geometry,xcolor}
\geometry{margin=2.5cm}
\newcommand{\code}[1]{\texttt{#1}}

\title{Internal Architecture and Type--Checking Algorithms of Contemporary Proof Assistants}
\author{}
\date{}

\begin{document}
\maketitle
\vspace{-2em} % no introduction, jump straight in

\section{Layered Architecture Overview}
Contemporary proof assistants adopt a stratified design that separates a \emph{tiny, trusted kernel} from progressively less–trusted outer layers.  Fig.~\ref{fig:layers} summarises the canonical stack used by Coq, Lean 4, Agda, Isabelle/HOL and HOL Light\cite{coq-arch,isabelle-manual,hollight-tut,lean4-survey,agda-note}.

\begin{figure}[h]
\centering
\begin{tabular}{|c|}
\hline
\textbf{User Interfaces} (VS Code, Proof General, CoqIDE, etc.) \\\hline
\textbf{Elaborator / Front-End}\\
\hline
\textbf{Tactic Engine \& Automation}\\
\hline
\textbf{Kernel (Type Checker + Inference Rules)}\\
\hline
\textbf{Logical Foundations (CIC, HOL, MLTT, …)}\\
\hline
\end{tabular}
\caption{Standard proof-assistant layer cake.}
\label{fig:layers}
\end{figure}

\section{Kernels in Practice}
\subsection{Size and Trusted Computing Base}
Each system’s kernel implements only the primitive inference rules of its object logic, plus definitional equality.  
Coq’s kernel (\(\approx\)6 kLoC OCaml) checks elaborated \emph{proof terms} for well-typedness in the Calculus of Inductive Constructions (CIC)\cite{coq-arch}.  
Lean 4’s C++ kernel is slightly larger (\(\approx\)9.5 kLoC) but mirrors the same minimalist philosophy\cite{lean4-survey}.  
HOL Light, in contrast, fits its entire kernel into \(\approx\)2.3 kLoC of OCaml by exploiting HOL’s small rule set\cite{hollight-tut}.

\subsection{De Bruijn Criterion}
All surveyed assistants satisfy the \emph{de Bruijn criterion}: every externally supplied proof is reduced to kernel-checked primitives, guaranteeing that only the kernel must be trusted\cite{coq-core}.  Isabelle achieves this by embedding object logics inside a generic meta-logic implemented in Standard ML\cite{isabelle-manual}.

\section{Type-Checking Algorithms}
\subsection{Bidirectional Checking}
Coq, Lean 4 and Agda implement \emph{bidirectional} algorithms that alternate between
\emph{checking} (\(\Gamma\vdash t:\!T\)) and \emph{inference} (\(\Gamma\vdash t\Rightarrow T\)) modes to localise where
conversion (\(\equiv\)) needs to be solved\cite{bidirectional-cic,tc-cic}.  
Lean 4 pushes most reductions eagerly (strong head β-normalisation) to minimise expensive definalisation at the leaves\cite{lean4-survey,lean4lean}.

\subsection{Conversion and Normalisation}
All kernels rely on decidable conversion: two terms are definitionally equal iff their
normal forms are syntactically identical under β (and sometimes ι, ζ, δ) reduction.  
Coq uses a weak-head reduction with on-the-fly unfolding of transparent constants;  
Agda employs \emph{proof-irrelevance annotations} to erase compile-time proofs before equality checking, thereby accelerating normalisation\cite{agda-note}.

\subsection{Universe Management}
Lean 4 and Coq feature cumulative universe hierarchies.  Lean’s algorithm stores
constraints in a union–find structure and relies on Tarjan’s algorithm for \(\le\)–closure\cite{lean4-survey}.  
Agda instead treats universe levels as first-class terms subject to unification, simplifying metaprogramming at the cost of heavier constraints\cite{agda-guide}.

\section{Term Representation}
\paragraph{Names.}  Coq, Lean and Agda all compile surface names to \emph{de Bruijn indices},
ensuring α-equivalent terms share a binary encoding\cite{coq-core,lean4-survey}.  
Isabelle and HOL Light keep explicit identifiers because their HOL core lacks binding-sensitive rules\cite{isabelle-manual,hollight-tut}.

\paragraph{Hash-Consing.}  Lean 4 hashes every node and maintains pointer equality to allow \(O(1)\) conversion checks on already-normalised sub-terms\cite{lean4-survey}.  HOL Light uses a related pointer-tagging trick for quick syntactic comparisons\cite{hollight-tut}.

\section{Elaboration Front-Ends}
The elaborator translates user syntax (with holes, overloading, implicit arguments) into fully explicit kernel terms.
Coq’s \code{Vernac} language feeds an OCaml elaborator that performs first-order unification followed by metavariable resolution\cite{coq-arch}.  
Lean 4’s elaborator is written partly in Lean itself and exploits reflective tactics; it is intentionally \emph{not} in the trusted base\cite{lean4lean}.  
Agda’s interactive mode allows partially written programs; its elaborator inserts `?`-holes and later solves them by constraint propagation\cite{agda-guide}.

\section{Automation and Tactics}
While tactics are outside the trusted core, their output is merely a proof term later verified by the kernel, preserving soundness\cite{coq-arch}.  
Isabelle/Isar uses a declarative proof language whose statements are compiled into kernel inferences\cite{isabelle-manual}.

\section{Persistence and Compilation}
Coq compiles verified libraries into \code{.vo} objects that cache the normalised term and universe constraints to accelerate later replay; Lean 4 analogously stores \code{.olean} files\cite{lean4-survey}.  
HOL Light relies on OCaml’s marshalled values, whereas Isabelle uses poly/ML heaps\cite{hollight-tut,isabelle-manual}.

\section{Verified Kernels}
Research projects like \emph{Candle}, a CakeML-verified re-implementation of HOL Light, show that entire kernels can be machine-checked down to machine code\cite{candle}.  MetaCoq and Lean4Lean pursue analogous self-verification for CIC and Lean respectively\cite{metacoq,bidirectional-cic,lean4lean}.

\section{Conclusion (omitted)}
% No conclusion per user request
\bibliographystyle{plain}
\begin{thebibliography}{10}

\bibitem{coq-arch}
C.~Paulin-Mohring.
\newblock \emph{Introduction to the Coq Proof Assistant} (Course notes, 2025). [1]

\bibitem{coq-core}
Coq development team.
\newblock \emph{Core Language Reference Manual}, version 8.18.0. [2]

\bibitem{isabelle-manual}
T.~Nipkow, L.~C. Paulson, et al.
\newblock \emph{Isabelle System Manual}. [3][4]

\bibitem{hollight-tut}
J.~Harrison.
\newblock \emph{HOL Light Tutorial}. [5]

\bibitem{lean4-survey}
X.~Tang.
\newblock A Comprehensive Survey of the Lean 4 Theorem Prover. \emph{arXiv 2501.18639}, 2025. [6]

\bibitem{lean4lean}
M.~Carneiro.
\newblock Lean4Lean: a Lean 4 kernel in Lean. GitHub repository, 2025. [22]

\bibitem{agda-note}
Q.~Pu.
\newblock A Note on the Agda Codebase—Type Checking \& Reflection. HackMD, 2023. [8]

\bibitem{agda-guide}
Agda team.
\newblock Quick Guide to Editing, Type Checking and Compiling Agda Code, 2022. [11]

\bibitem{bidirectional-cic}
M.~Lennon-Bertrand.
\newblock Complete Bidirectional Typing for the Calculus of Inductive Constructions. \emph{ITP 2021}. [9]

\bibitem{tc-cic}
R.~Pollack.
\newblock Type Checking in Pure Type Systems. Technical Report, 1994. [12]

\bibitem{candle}
A.~Kumar et al.
\newblock Candle: A Verified Implementation of HOL Light. \emph{ITP 2022}. [10]

\bibitem{metacoq}
A.~Anand et al.
\newblock MetaCoq project: certifying Coq’s metatheory. Project page, 2018. [9]

\end{thebibliography}
\end{document}

### 2025-07-29 Mainpaper/paper.tex
test

@@
 
-% Finally rigidify the basics refer to the original paper
-
-% Update based on knowledge.....
+        .
 
 \section{General Architecture of Proof Assistants}
+The proof assistants follow a general architecture that adopts a design that seperates tiny,trusted kernel from progressively less trusted outer layer[]
+\begin{figure}[h]
+\centering
+\begin{tabular}{|c|}
+\hline
+\textbf{User Interfaces} (VS Code, Proof General, CoqIDE, etc.) \\\hline
+\textbf{Elaborator / Front-End}\\
+\hline
+\textbf{Tactic Engine \& Automation}\\
+\hline
+\textbf{Kernel (Type Checker + Inference Rules)}\\
+\hline
+\textbf{Logical Foundations (CIC, HOL, MLTT, …)}\\
+\hline
+\end{tabular}
+\caption{Standard proof-assistant layer cake.}
+\label{fig:layers}
+\end{figure}
 
+\subsection{The kernel} 
+At the heart of any proof assistant lies the kernel, which is a minimal and formally verified component responsible for enforcing the programming logical rules.
+kernel implements the foundational calculus or logic and validates every proof
+step by checking derivations against these rules.
+The kernel does't trust the computing base and checks for every proof , which ensures that the entire system depends on this critical component [Gordon, 1993][Paulson, 1987][The Coq Development Team, 2025].
+\subsection{Tatic Engine}
+Above the kernel layer is the tactic engine that assists users in constructing proofs interactively or automatically.
+Tactics provide programmable strategies to decompose proof goals into simpler subgoals which can be easier to prove. 
+Tactic implementations can become complex, but since any output they generate must be verified through the
+kernel before acceptance,system integrity is maintained [Nipkow et al., 2002][Delahaye, 2000].
+
+\subsection{Formal proof language}
+Proof assistants offer a dedicated programming language designed to express definitions, statements and proofs. These languages are typically Dependently typed,
+and allows implicit argument . they are highly readable and rigorous. Integration with tactic scripting enables a smooth blend of automated and manual proof development [Barras et al.,
+1997][Agda Team, 2023][Lean Development Group, 2024].
+
+\subsection{Type checker and inference engine}
+The type checker enforces logical consistency by validating the written expression , checking if it meets the system's typing
+rules ,importantly preventing paradoxical constructs. Proof assistant often use bi-directional type checking and other 
+complex inference algorithms.
+
+\subsection{User Interface} 
+Effective user interfaces (UIs) enable seamless interaction with the proof assistant. 
+Modern UIs include integrated development environments (IDEs) or editor plugins that provide syntax highlighting, error diagnostics, real-time proof state visualization,
+and automated tactic suggestion. Such interfaces significantly enhance productivity and make formal methods more accessible to a wider audience [CoqIDE, 2025][Isabelle/jEdit, 2023].
+\subsection{formal libraries}
+Proof assistants come with extensive, rigorously verified libraries that
+encapsulate fundamental theories (e.g., arithmetic, set theory, algebra), supporting rapid development of
+complex formalizations without reinventing foundational results.
+These libraries grow continuously and provide a shared repository of reusable components vital for large-scale verification projects [Coq Standard Library, 2025][Mathlib, 2025][Isabelle Archive of Formal Proofs].
 \section{Upon Some Proof Assistants and Comparative Study}
-\subsection{Coq}
+\subsection{ Agda}
+Agda  is a Dependently typed functional programming language and proof assistant. Extension of Martin-Lof's Type theory. (agda documentation) 
+In Dependently typed programming language there is no clear distinction between types and values , and types can depend on arbitary values and may appear as results of ordinary functions. 
+This feature of Dependently typed programs allows us to encode properties of values as type of proofs of that property, allowing us to use Dependent type programming as logic.
+(Dependently Typed Programming in Agda)
+
+Every agda file has a top level module which corresponds to name of the file and the main programs goes inside the top level module.[]
+Another key component of agda is pattern matching. Pattern matching is a mechanism for analysing the structure of values. When inductively defined  datatypes are introduced the
+pattern matching becomes even more powerful as patten matching on one value can give the information about another value (thesis)
+
+\subsection{ Type checking algorithm}
+Type checking algorithm in agda is done by first taking a sequence of patterns and checking against telescope  
+\subsection{Rocq}
+Rocq formally known as coq is another proof assistant or interactive theorem prover and lets you formalize mathematical proofs. becoming famous for the Formalization of four color theorem  .
+Rocq is vastly different from agda in terms of proofs, While Agda is type theory based direct construction proof assistant Rocq on the other hand is tatic based. 
+In Rocq users generate proofs by entering a series of tatics, which constitute steps of proofs
+\subsubsection{Tatics}
+    Mainly there are two ways of generating proofs forward reasoning and backward reasoning  \\ 
+    In forward reasoning the simpler and smaller statements are proved and constructed to bigger statements and theorem, combining to prove the theorem at the end \\
+    In Backward reasoning proofs begins with theorem statement and gradually transformed into sub-statement or goals which are proven. 
+    Rcoq and Tatics heavily relies on Backward reasoning models.
+    A tatic by itself may fully prove a goal or divide into sub goals to prove and most tatic in rocq require certain condiditon or elements to reduce a goal.
+    some tatic are automatic and can solve complex goals\\ (doc website)
+    \subsubsection{kernel}
+
+    Rocq's kernel is based on PCUIC , a Dependently typed lambda calculus extended with various other rules and features.
+Rocq's core kernel converts all the notations and implicit arguments into core language which is the calculus of inductive construction.
+    the kernel wil verify the proof term built by the tatics . This type of seperation is called the "de Bruijn criterion" and makes it so that only the smaller components
+    like the kernel is trusted and the entire program is checked by the kernel.
+
+\subsubsection{Type checking}
+        In Rocq the syntaxes and typing rules are specified so that they remain mathematicaly precise and machine verifiable. For that the type system is first described declaratively, then refined to
+        algorithmic specification.
+Rocq introduces a bi-directional infer and check type system that guides implementation.The checker is implemented in coq itsel.
+Much of type checking relies on deciding if two tpes are equivalent by Conversion or if one is the subtype of
+        other by cumulativity which is very challenging because these relations depend on reduction.
+The type checker is constructed as bi-directional algorithm and to ensure soundess the checker carries proof carrying code and returns the proof along with
+        the type checking decision.
+Every critical theoretic properity need for type checking is proved or assumed axiomatically to be true.
+(Correct and Complete Type Checking and Certified Erasure for Coq, in Coq)
 
 \subsection{Lean} 
+Lean theorem prover is a proof assistant developed by leonardo di-moura at microsoft in 2013. one way to prove theorem in lean is to 
+build from assersion which follows the calculus of construction. Lean treats any two element say 
+$ t1 \ t2: p $ to be equal, which is known as proof irrelevance. And allows for proposition as types logic. calculational proofs can also be written in lean and starts 
+with the word calc.
 
-\subsection{Isabelle}
-
-\subsection{Agda}
-
-\section{ Agda}
+Another way Like Rocq Lean can also relie on tatics. Tatics are useful for construction and destruction of data and there are many built in tatics avaible in lean. One of the major advantage of lean 
+is that you can mix term-style and tatic-style proof writing and pass freely between the two. (lean-lang.org) Lean allows for unsafe declaration but cannot be exported
+Lean export file
 
 \section{Formalization of Some Proofs}
+Every agda file begins with a top level module declaration which corresponds to the name of the file, 
+and the following code is written under the, module. 
+eg: 
+\begin{verbatim}
+        module hello where
+\end{verbatim}
+
+An example of an implication in agda could be 
+\begin{verbatim}
+        module implication_module
+        an-implication: {A B C : set} -> (A->B) -> (B -> C) -> A -> C 
+        an-implication g g' f = g'(g x)
+\end{verbatim}
 

### 2025-07-29 Mainpaper/paper.tex
Added Proofs to paper, note that de morgan aint rendered Need to fix unicode rendering

@@
 
+\clearpage
+\section*{}
+\addcontentsline{toc}{section}{\textbf{ABSTRACT}}
+
 \begin{abstract}
+This report explores the theoretical foundations and working principles of proof assistants, with a focus on Agda.
+It introduces key concepts from logic, type theory, and the Curry–Howard correspondence that underlie formal verification.
+We discuss how proofs are treated as programs and propositions as types in dependently typed systems.
+A comparative study of proof assistants such as Agda, Coq (Rocq), and Lean is included, highlighting their kernel architectures and type-checking algorithms.
+Finally, we demonstrate formalization of selected logical proofs in Agda, reflecting on challenges and limitations encountered during implementation.
 \end{abstract}
+\clearpage
 
@@
 \section{Introduction}
+
+In 1998 Thomas C. Hales announced that he had proved the kepler conjecture
+(Solane,1998) However the peer review took over 4 years especially due to the proof
+being incredibely difficult to check, with over a dozen of mathematician to refree
+the proof and although they were 99% certain it was not untill the proof became
+formalized that they were able to completely validate the correctness of the proof,
+which according to Thomas Hales, would have taken 20 person-year of manual work.
+[Szpiro, G, 2003]
+
+Proof assistants helps with formalization of mathematical proofs in computer which
+enables for their verification digitally. Some exhaustive proofs may contain large
+number of cases which is not feasible to write and verify manually and requires
+computer assitance.
+Such Proof assistants (systems) work by treating proofs as computational objects
+and checking them using strict logical rules. To enable this, proofs must be writ-
+ten in a formal language that removes ambiguity and allows machines to interpret
+them accurately. Underlying this process is a theoretical framework—often type
+theory—that defines how propositions and proofs relate to each other.
+This method bridges programming and logic: writing a proof resembles writing a
+program, and verifying it is like type-checking. By interpreting logic computation-
+ally, proof assistants ensure that every proof is exact, complete, and verifiable by
+a machine. This project invloves investigating this process.
+
+\subsection{Objectives}
+\begin{enumerate}
+    \item To explore the correspondence between Proofs and Programs, Type Theory and Intuitionstic Logic.
+    \item To investigate a current exsisting proof assistants (Agda) and identify the mathematics behind it.
+    \item To formalize some selected proofs in Agda and demonstrate verification process.
+\end{enumerate}
+
+\subsection{Limitations}
+\begin{enumerate}
+    \item This study fails to rigorously present the Type Checking Algorithm.
+    \item Avoids formalization of complicated proofs.
+\end{enumerate}
 
@@
 Then to write a  program (equivalent to proof) of the proposition $A \to B \to A$, we need to construct a function that takes term of type A , term of type B and returns term of type A, that is $\lambda a{:}A.\lambda b{:}B.t{:}A $ which proceeds as
-
+\begin{proposition}
+    $A \to B \to A$
    \begin{prooftree}
@@
 \end{prooftree}
+\end{proposition}
 
@@
 The kernel does't trust the computing base and checks for every proof , which ensures that the entire system depends on this critical component [Gordon, 1993][Paulson, 1987][The Coq Development Team, 2025].
-\subsection{Tatic Engine}
+\subsection{Tactic Engine}
 Above the kernel layer is the tactic engine that assists users in constructing proofs interactively or automatically.
@@
 
-\subsection{Formal proof language}
+\subsection{Formal Proof language}
 Proof assistants offer a dedicated programming language designed to express definitions, statements and proofs. These languages are typically Dependently typed,
@@
 and automated tactic suggestion. Such interfaces significantly enhance productivity and make formal methods more accessible to a wider audience [CoqIDE, 2025][Isabelle/jEdit, 2023].
-\subsection{formal libraries}
+\subsection{Formal Libraries}
 Proof assistants come with extensive, rigorously verified libraries that
@@
 These libraries grow continuously and provide a shared repository of reusable components vital for large-scale verification projects [Coq Standard Library, 2025][Mathlib, 2025][Isabelle Archive of Formal Proofs].
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.7\textwidth]{fig/ui.png}
+
+\end{figure}
+\cite{Geuvers20091}
+
 \section{Upon Some Proof Assistants and Comparative Study}
@@
 
-\subsection{ Type checking algorithm}
-Type checking algorithm in agda is done by first taking a sequence of patterns and checking against telescope  
+\subsubsection {Kernel}
+Agda's Kernel is written in Haskell.
+With no seperate tactic layer, the kernel is directly involved in all proof construction making it minimalistic yet more rigid.
+It tightly integrates normalization, which is a direct application of the Normalization Theorem in \cite{martinlof1972intuitionistic_theory}.
+
+\subsubsection{ Type Checking algorithm}
+Agda uses a bidirectional type checking algorithm grounded in Martin-Löf Type Theory, emphasizing manual proof construction and transparency.
+The checker alternates between inferring types for expressions and checking expressions against known types, supported by normalization by evaluation (NbE) to verify definitional equality.
+Unlike Lean or Coq, Agda offers minimal automation—there are no tactics, typeclass resolution, or automatic proof search.
+This makes it more predictable but also more labor-intensive.
+In contrast, Lean and Coq automate many aspects of type inference and proof construction via elaborators and tactics, with Coq relying on a trusted kernel and Lean integrating extensible unification and metaprogramming.
+Agda prioritizes explicitness and formal clarity, making it ideal for foundational studies in type theory.
+\cite{depPUlf}
 \subsection{Rocq}
-Rocq formally known as coq is another proof assistant or interactive theorem prover and lets you formalize mathematical proofs. becoming famous for the Formalization of four color theorem  .
-Rocq is vastly different from agda in terms of proofs, While Agda is type theory based direct construction proof assistant Rocq on the other hand is tatic based. 
-In Rocq users generate proofs by entering a series of tatics, which constitute steps of proofs
-\subsubsection{Tatics}
+Rocq formerly known as coq is another proof assistant or interactive theorem prover and lets you formalize mathematical proofs. becoming famous for the Formalization of four color theorem  .
+Rocq is vastly different from agda in terms of proofs, While Agda is type theory based direct construction proof assistant Rocq on the other hand is tactic based. 
+In Rocq users generate proofs by entering a series of tactics, which constitute steps of proofs.
+\subsubsection{Tactics}
     Mainly there are two ways of generating proofs forward reasoning and backward reasoning  \\ 
@@
     In Backward reasoning proofs begins with theorem statement and gradually transformed into sub-statement or goals which are proven. 
-    Rcoq and Tatics heavily relies on Backward reasoning models.
-    A tatic by itself may fully prove a goal or divide into sub goals to prove and most tatic in rocq require certain condiditon or elements to reduce a goal.
-    some tatic are automatic and can solve complex goals\\ (doc website)
-    \subsubsection{kernel}
+    Rcoq and Tactics heavily relies on Backward reasoning models.
+    A tactic by itself may fully prove a goal or divide into sub goals to prove and most tactic in rocq require certain condition or elements to reduce a goal.
+    some tactic are automatic and can solve complex goals\\ (doc website)
+    \subsubsection{Kernel}
 
-    Rocq's kernel is based on PCUIC , a Dependently typed lambda calculus extended with various other rules and features.
+    Rocq's kernel is based on Predicative Calculus of Cumulative Inductive Constructions(PCUIC) , a Dependently typed lambda calculus extended with various other rules and features.
 Rocq's core kernel converts all the notations and implicit arguments into core language which is the calculus of inductive construction.
-    the kernel wil verify the proof term built by the tatics . This type of seperation is called the "de Bruijn criterion" and makes it so that only the smaller components
+    the kernel wil verify the proof term built by the tactics . This type of seperation is called the "de Bruijn criterion" and makes it so that only the smaller components
     like the kernel is trusted and the entire program is checked by the kernel.
+Furthermore, its written in Coq itself (extracted to OCaml) i.e a Kernel being a program itself, its correctness is verified within it.
 
-\subsubsection{Type checking}
+\subsubsection{Type checking Algorithm }
         In Rocq the syntaxes and typing rules are specified so that they remain mathematicaly precise and machine verifiable. For that the type system is first described declaratively, then refined to
@@
 Much of type checking relies on deciding if two tpes are equivalent by Conversion or if one is the subtype of
-        other by cumulativity which is very challenging because these relations depend on reduction.
+        other by cumulatively which is very challenging because these relations depend on reduction.
 The type checker is constructed as bi-directional algorithm and to ensure soundess the checker carries proof carrying code and returns the proof along with
@@
 Every critical theoretic properity need for type checking is proved or assumed axiomatically to be true.
+The bidirectional type checking employed by Rocq is weak compared to Agda but it relies heavily on conversion.
 (Correct and Complete Type Checking and Certified Erasure for Coq, in Coq)
@@
 \subsection{Lean} 
-Lean theorem prover is a proof assistant developed by leonardo di-moura at microsoft in 2013. one way to prove theorem in lean is to 
-build from assersion which follows the calculus of construction. Lean treats any two element say 
+Lean theorem prover is a proof assistant developed by Leonardo Di-Moura at Microsoft in 2013. one way to prove theorem in lean is to 
+build from assertion which follows the calculus of construction. Lean treats any two element say 
 $ t1 \ t2: p $ to be equal, which is known as proof irrelevance. And allows for proposition as types logic. calculational proofs can also be written in lean and starts 
@@
 
-Another way Like Rocq Lean can also relie on tatics. Tatics are useful for construction and destruction of data and there are many built in tatics avaible in lean. One of the major advantage of lean 
-is that you can mix term-style and tatic-style proof writing and pass freely between the two. (lean-lang.org) Lean allows for unsafe declaration but cannot be exported
-Lean export file
+Another way Like Rocq Lean can also relie on tactics.
+Tactics are useful for construction and destruction of data and there are many built in tactics avaible in lean.
+One of the major advantage of lean
+is that you can mix term-style and tactic-style proof writing and pass freely between the two. 
+The ease of working with Lean is responsible for its popularity against other Proof Assistants.
+% (lean-lang.org) Lean allows for unsafe declaration but cannot be exported
+% Lean export file 
+\subsubsection{Kernel}
+Lean's kernel is written in C++ (Lean 3) and C (Lean 4).
+The kernel much like Coq is based of Calculus of Inductive Constructions and employs conversion checking.
+Though its elaboration and tactic system is large, it is relatively less rigid but improves usability.
+Its performance based and offers flexibility.
+Lean's kernel is implemented in two layers. The First layer contains the
+type checker and APIs for creating and manipulating terms, declarations, and
+the environment and the second layer provides additional components such as inductive families and quotient types. When the kernel is
+instantiated, one selects which of these components should be used.
+\cite{Lean}
+\subsubsection{Type Checking Algorithm}
+Likewise,the Bidirectional algorithm is applied here too. But the strength of Lean lies in its smart elaboration. Lean offers coercion, backtracking and overloading.
+
+\subsection{The Bidirectional Type Checking Algorithm }
+
+At its core, this algorithm works on two modes 
+\begin{itemize}
+    \item Checking Mode $(\Gamma \vdash M \leftarrow A )$ Given the term M and an expectied type A , the task is to verify that M has type A.
+Its done when the type is known or can be known then propagated from the context.
+    \item Synthesis (Inference) Mode  ($\Gamma \vdash M \rightarrow A$ ) When the type is known, the job is to infer it.
+That is synthesize unique type A for the term M.
+Since the type is not available from the context $\Gamma$ , it needs to be determined upwards.
+
+\end{itemize}
+The checking and infering in two directions (from or to context $\Gamma$) led it to be called Bidirectional .
+\\
+
+\clearpage
 
 \section{Formalization of Some Proofs}
-Every agda file begins with a top level module declaration which corresponds to the name of the file, 
-and the following code is written under the, module. 
-eg: 
+
+\subsection{Defining Natural Numbers, Equality and Proof of Assocaitive and Commutative Properties}
 \begin{verbatim}
-        module hello where
+
+data N : Set where
+    zero : N 
+    suc : N -> N
+
+{-# BUILTIN NATURAL N #-}
+
+--Equality
+data _==_ { A : Set } ( x : A ) : A -> Set where 
+    refl : x == x
+
+cong : {A B : Set } ( f : A -> B ) { x y : A}
+    -> x == y
+    -> f x == f y
+cong f refl = refl 
+sym : forall { A : Set } { x y : A}
+    -> x == y
+    -> y == x 
+sym refl = refl
+
+trans : { A : Set } { x y z : A}
+    -> x == y
+    -> y == z
+    -> x == z
+
+trans refl refl = refl
+
+{-# BUILTIN EQUALITY _==_ #-}
+
+infix 4 _==_
+
+--ADDITION
+
+_+_ : N -> N -> N 
+zero + x = x
+suc x + y = suc(x + y)
+
+--ZERO
+zero-+ : ( a : N ) -> ( ( zero + a ) == a )
+zero-+ a = refl
+
++-zero : ( a : N ) -> ( ( a + zero ) == a )
++-zero zero = refl
++-zero (suc a) = cong suc (+-zero a)
+
+--LEMMA FOR COMMUTATIVITY
++-suc : forall x y -> x + suc y == suc (x + y)
++-suc zero y = refl
++-suc (suc x) y = cong suc (+-suc x y)
+
+--ASSOCIATIVITY
+assoc-+ : forall x y z -> (x + y) + z == x + (y + z)
+assoc-+ zero y z = refl
+assoc-+ (suc x) y z = cong suc (assoc-+ x y z)
+
+--COMMUTATIVITY
++-comm : ( a b : N) -> ( ( a + b ) == ( b + a ) )
++-comm a zero = +-zero a 
++-comm a (suc b) = trans (+-suc a b) (cong suc (+-comm a b))
+
 \end{verbatim}
 
-An example of an implication in agda could be 
+\subsection{Formalizing Proposition 2.3 and 2.4 }
 \begin{verbatim}
-        module implication_module
-        an-implication: {A B C : set} -> (A->B) -> (B -> C) -> A -> C 
-        an-implication g g' f = g'(g x)
+module props where
+
+open import Agda.Primitive using (Level; lzero)
+
+open import Data.Product using (Sigma; _,_)
+
+variable
+  l : Level
+  X : Set l
+  A B : X → Set l
+  s : X
+
+prop23 : (forall x → A x → B x) → (s : X) → A s → Sigma X B
+prop23 allImp s aS = s , allImp s aS
+--Replace sigma with greek alphabet while loading.
+
+prop24 : forall {A B : Set} -> A -> B -> A
+prop24 a b = a
 \end{verbatim}
+
+\subsection{DeMorgan's Law}
+\begin{verbatim}
+    open import Agda.Primitive using (Level; lzero)
+open import Data.Product using (_×_; _,_)
+open import Data.Sum using (_⊎_; inj₁; inj₂)
+open import Relation.Nullary using (Dec; yes; no)
+open import Data.Empty using (⊥; ⊥-elim)
+open import Relation.Nullary.Negation using (¬_)
+
+--One Direction
+deMorganOneWay : ∀ {ℓ} {P Q : Set ℓ} → (¬ P) ⊎ (¬ Q) → ¬ (P × Q)
+deMorganOneWay (inj₁ np) (p , q) = np p
+deMorganOneWay (inj₂ nq) (p , q) = nq q
+
+--Converse, Requires Non Constructive Assumptions
+deMorganOtherWay :
+  ∀ {ℓ} {P Q : Set ℓ}
+  → Dec P
+  → Dec Q
+  → ¬ (P × Q)
+  → (¬ P) ⊎ (¬ Q)
+deMorganOtherWay (yes p) _ notPQ = inj₂ (λ q → notPQ (p , q))
+deMorganOtherWay (no np) (yes q) _ = inj₁ np
+deMorganOtherWay (no np) (no nq) _ = inj₁ np  -- or inj₂ nq
+
+\end{verbatim}
+\subsection{Euclid's Theorem of Primes }
+
+\subsection{Euclid's Algorithm's Proof}
+
+\subsection{Attempt to Constructive Reals}
 

### 2025-08-04 Mainpaper/paper.tex
minor fixes in unicode notation and declaration
cover page , ack page and minor changes

@@
 \usepackage{hyperref}
+\usepackage{graphicx}
 \usepackage{cite}
@@
 \usepackage{titlesec}
+\usepackage{newunicodechar}
 \usepackage{tikz}
+\DeclareUnicodeCharacter{2200}{\ensuremath{\forall}}    % ∀
+\DeclareUnicodeCharacter{2113}{\ensuremath{\ell}}       % ℓ
+\DeclareUnicodeCharacter{2192}{\ensuremath{\rightarrow}}% →
+\DeclareUnicodeCharacter{00D7}{\ensuremath{\times}}     % ×
+\DeclareUnicodeCharacter{228E}{\ensuremath{\uplus}}   % ⊎
+\DeclareUnicodeCharacter{22A5}{\ensuremath{\bot}}       % ⊥
+\DeclareUnicodeCharacter{00AC}{\ensuremath{\neg}}       % ¬
+\DeclareUnicodeCharacter{2081}{\textsubscript{1}}       % ₁
+\DeclareUnicodeCharacter{2082}{\textsubscript{2}}       % ₂
+\DeclareUnicodeCharacter{03BB}{\ensuremath{\lambda}}    % λ
+\DeclareUnicodeCharacter{03A3}{\ensuremath{\Sigma}}   % Σ
+\geometry{
+  left=1.2in,
+  right=1.2in,
+  top=1.1in,
+  bottom=1.1in
+}
 \usetikzlibrary{arrows.meta, positioning}
@@
 }
-%mention supervisor
 \date{\today}
@@
 
-\maketitle
+\begin{titlepage}
+    \begin{center}
+        % Title
+        \Large\bfseries
+        Working Principles of Proof Assistants and Formalization of some proofs in Agda\\[2cm]
+
+        \normalsize
+        SECOND YEAR PROJECT REPORT\\[2ex]
+
+        \small
+        SUBMITTED IN PARTIAL FULFILLMENT OF THE REQUIREMENTS FOR\\
+        THE DEGREE OF MSC. IN COMPUTATIONAL MATHEMATICS\\[2cm]
+
+        \normalsize
+        BY\\[2ex]
+        \textbf{Ashwot Acharya, Bishesh Bohora, Supreme Chaudhary}\\[2cm]
+
+        \includegraphics[width=4cm]{logo.png}\\[2cm]
+        % Replace with your logo file. Place 'ku_logo.png' in the same folder!
+
+        DEPARTMENT OF MATHEMATICS\\
+        KATHMANDU UNIVERSITY\\
+        DHULIKHEL, NEPAL\\[2ex]
+
+        \vfill
+    \end{center}
+\end{titlepage}
+
+\thispagestyle{empty}
+
+\begin{center}
+    \textbf{\LARGE CERTIFICATION}
+\end{center}
+
+\vspace{1em}
+
+This project entitled ``\textbf{Working Principles of Proof Assistants and Formalization of some proofs in Agda}'' is carried out under our supervision for the specified entire period satisfactorily and is hereby certified as a work done by the following students:
+\begin{enumerate}
+    \item Ashwot Acharya
+    \item Bishesh Bohora 
+    \item Supreme Chaudhary
+\end{enumerate}
+in partial fulfillment of the requirements for the degree of B.Sc. in Computational Mathematics, Department of Mathematics, Kathmandu University, Dhulikhel, Nepal.
+
+\vspace{3em}
+
+\noindent
+\begin{minipage}{0.45\textwidth}
+\rule{6cm}{0.5pt}\\
+\textbf{M.r K.B Manandhar }\\
+Assistant Professor\\
+Department of Mathematics,\\
+School of Science\\ Kathmandu University,\\
+Dhulikhel, Kavre, Nepal\\
+Date: August, 2025
+
+\end{minipage}
+
+\vspace{2.5em}
+
+\noindent
+\\ \\ 
+\textbf{APPROVED BY:}\\
+I hereby declare that the candidate qualifies to submit this report of \textbf{Working Principles of Proof Assistants and Formalization of some proofs in Agda} to the Department of Mathematics.
+\\
+\vspace{1.7em}
+\\
+\noindent
+\rule{6cm}{0.5pt}\\ \\
+\textbf{Head of Department }\\
+Department of Mathematics\\
+School of Science\\
+Kathmandu University\\
+Date: August, 2025
+
+\vspace{2em}
+
+\begin{center}
+   {\small ii}
+\end{center}
+\clearpage
+
+\addcontentsline{toc}{section}{\textbf{ABSTRACT}}  
+\clearpage
+
+\begin{center}
+    \textbf{\LARGE ACKNOWLEDGMENTS}
+\end{center}
+    First and foremost, we would like to thank our supervisor Mr K.B Manandhar sir for guiding us through this research project.
+Their feedback and suggestions helped us overcome numerous challenges and shaped
+our understanding of the subject matter.
 
 \clearpage
-\section*{}
-\addcontentsline{toc}{section}{\textbf{ABSTRACT}}
 
@@
 
-prop23 : (forall x → A x → B x) → (s : X) → A s → Sigma X B
+prop23 : (∀ x → A x → B x) → (s : X) → A s → Σ X B
 prop23 allImp s aS = s , allImp s aS
---Replace sigma with greek alphabet while loading.
 
-prop24 : forall {A B : Set} -> A -> B -> A
+prop24 : ∀ {A B : Set} -> A -> B -> A
 prop24 a b = a

### 2025-08-05 Mainpaper/paper.tex
minor change
minor changes added some missed references

@@
         SUBMITTED IN PARTIAL FULFILLMENT OF THE REQUIREMENTS FOR\\
-        THE DEGREE OF MSC. IN COMPUTATIONAL MATHEMATICS\\[2cm]
+        THE DEGREE OF BSC. IN COMPUTATIONAL MATHEMATICS\\[2cm]
 
@@
 Rocq is vastly different from agda in terms of proofs, While Agda is type theory based direct construction proof assistant Rocq on the other hand is tactic based. 
-In Rocq users generate proofs by entering a series of tactics, which constitute steps of proofs.
+In Rocq users generate proofs by entering a series of tactics, which constitute steps of proofs. \cite{rocq_manual}
 \subsubsection{Tactics}
@@
     A tactic by itself may fully prove a goal or divide into sub goals to prove and most tactic in rocq require certain condition or elements to reduce a goal.
-    some tactic are automatic and can solve complex goals\\ (doc website)
+    some tactic are automatic and can solve complex goals\\ \cite{rocq_tactics}
     \subsubsection{Kernel}
@@
 Furthermore, its written in Coq itself (extracted to OCaml) i.e a Kernel being a program itself, its correctness is verified within it.
+\cite{rocq_core_language}
 
@@
 The bidirectional type checking employed by Rocq is weak compared to Agda but it relies heavily on conversion.
-(Correct and Complete Type Checking and Certified Erasure for Coq, in Coq)
+(Correct and Complete Type Checking and Certified Erasure for Coq, in Coq) \cite{type_checking}
 
@@
 $ t1 \ t2: p $ to be equal, which is known as proof irrelevance. And allows for proposition as types logic. calculational proofs can also be written in lean and starts 
-with the word calc.
+with the word calc.  \cite{lean_functional_programming}
 
@@
 is that you can mix term-style and tactic-style proof writing and pass freely between the two. 
-The ease of working with Lean is responsible for its popularity against other Proof Assistants.
-% (lean-lang.org) Lean allows for unsafe declaration but cannot be exported
-% Lean export file 
+The ease of working with Lean is responsible for its popularity against other Proof Assistants.\cite{lean_theorem_proving}
+
 \subsubsection{Kernel}

### 2025-08-05 project-defense/Presentation.tex
small changes tried some euclid codes

--- 2025-06-01 Presentation/Presentation.tex
@@
 \usepackage{cite}
+\usepackage{booktabs}
 
 \title{Working Principles of Proof Assistants and Formalization of Some Proofs in Agda}
-\author{Ashwot Acharya, Bishesh Bohora, Supreme Chaudary}
+\author{Ashwot Acharya, Bishesh Bohora, Supreme Chaudary \\Supervisior: K.B Manandhar}
+
 \institute{Kathmandu University}
@@
 \begin{frame}{Problem Statement}
-    As the title suggests, our project will revolve around exploration of theoritical fondations behind Proof Assistants and practice them.
+    As the title suggests, our project will revolve around exploration of theoretical foundation behind Proof Assistants and practice them.
 
@@
   \item Discovered type theory through internet memes on category theory.
-  \item Fascinated by the Four Colour Theorem and its computer-assisted proof.
+  \item Fascinated by the Four Color Theorem and its computer-assisted proof.
   \item The rise of AI raised the question: \textit{"How do computers understand reasoning?"}
@@
 \end{frame}
-\begin{frame}{Proof Assistants}
+\begin{frame}{Proof Assistant}
         In computer science and mathematical logic, a proof assistant or interactive theorem prover is a software tool to assist with the development of formal proofs by human–machine collaboration.\cite{wikipedia_proof_assistant}
@@
 
-\section{Glance At Theoritical Foundations}
+\section{Glance At Theoretical Foundations}
 \begin{frame}{Type Theory}
@@
 
-  \item In proof assistants, Core logic is based on typed lambda calculus.
+  \item In proof assistant, Core logic is based on typed lambda calculus.
       \item Tools like Coq and Agda embed functional languages with type theory.
@@
 \begin{frame}{Methodology}
-
+    \begin{itemize}
+      \item Investigating the Coq and the agda proof assistant
+      \item Assessing the logic behind each of these proof assistants
+      \item Investigating of formalization of proofs in Agda
+      \item Dissection of verification process
+    \end{itemize}
 \end{frame}
 \begin{frame}{Agda}
-Agda is a functional programming language with dependent types. It is based on Martin Löf Type Theory. And most importantly it is a proof assistants.
+Agda is a functional programming language with dependent types. It is based on Martin Löf Type Theory. And most importantly it is a proof assistant.
 \cite{inproceedings}
@@
 \begin{frame}{Why Agda?}
+  \begin{itemize}
+    \item Dependently Typed programming language
+    \item Fully embraces the Curry–Howard isomorphism
+    \item More comprehensible code
+    \item Active Community
+  \end{itemize}
 \end{frame}
@@
   \begin{center}
-    \begin{tabular}{|c|c|}
+    \begin{tabular}{p{1cm} p{8cm}}
         \hline
@@
     \item Improved understanding of proofs, logic and programming.
-    \item Experience of Using A Proof Assistant and incorporating them in regular studies.
     \item New Formalization of Proofs of various fields in Agda 

### 2025-08-05 Mainpaper/paper.tex
small changes tried some euclid codes
workarounds

--- 2025-08-04 Mainpaper/paper.tex
@@
 
+\begin{itemize}
+    \item In 5.1, we define natural numbers and their equality (When can we say two numbers are equal) from scratch.
+Defining the basic objects in every proof we write is tedious, therefore we use Agda's Standard Library and use pre defined and some pre-proven results to work further.
+    \item In 5.3, the forward direction was do able with constructive principles. But for the other way, we used law of excluded middle by importing Dec from Relation.Nullary.
+    \item 
+\end{itemize}
+
 % ======== References =========

### 2025-08-05 Mainpaper/paper.tex

--- 2025-08-05 Mainpaper/paper.tex
@@
 
-% ======== Packages =========
 \usepackage[utf8]{inputenc}

### 2025-08-05 final-presentation/final.tex
final presentation kYOUR

\documentclass[aspectratio=169, compress, 12pt]{beamer}

\usepackage[english]{babel}                 %%%%%%%%%%%%%%%%%% (EDIT LANGUAGE)

\usepackage{Custom}                         %%%%%%%%%%%%%%%%%%    (EDIT NAMES)
\usepackage{Custom-2}                         %%%%%%%%%%%%%%%%%%    (EDIT COLOURS)
\usepackage{AuxiliaryFiles/AuxiliaryFiles}  %%%%%%%%%%%%%    (DO NOT EDIT)

\begin{document}
\TitlePage        
\SectionPage    
\SubsectionPage    %%%% PAGE AT THE BEGINNIN OF A SECTION - IF NOT NEEDED, COMMENT IT
\ProgressBar       %%%% PROGRESS BAR ON THE RIGHT MARGIN - IF NOT NEEDED, COMMENT IT
\PageNumbering     %%%% SLIDES NUMBERING - IF NOT NEEDED, COMMENT IT

\input{AuxiliaryFiles/SettingBackup}

\input{AuxiliaryFiles/Thanks}  
\end{document}

### 2025-08-05 final-presentation/final.tex
slides edited upto comparative study

@@
 
+\section{Proof Assistants} 
+\subsection*{What are proof assistant}
+\begin{frame}{What are proof assistant}
+    proof assistant, are software more specifically a type of programming language thats allows us to 
+    formalize mathematical proofs in computer for digital verification. 
+\end{frame}
+\subsection*{Why digital verification is needed?}
+\begin{frame}{Need of digital verification}
+    \begin{itemize}
+        \item faster computation for complex problems
+        \item many exceptional cases can be explored which would take mathematicians  long time \\
+           \indent  \footnotesize ex:   The Kepler Conjecture's proof , which was so complex that verifying it manually would take 20 person-years, but proof assistants made this verification feasible and fast.
+    \end{itemize}
+\end{frame}
+\section{Logical foundation}
+\begin{frame}
+    \begin{itemize}
+    \item Based on logic (natural deduction, intuitionistic logic), $\lambda$-calculus, and type theory.
+    \item Curry–Howard Correspondence:
+    \begin{itemize}
+        \item Propositions $\leftrightarrow$ Types
+        \item Proofs $\leftrightarrow$ Programs
+    \end{itemize}
+    \item Dependently Typed Languages: Types can depend on values, enabling encoding of properties and proofs.
+\end{itemize}
+\end{frame}
+\subsection*{Naturl deduction}
+\begin{frame}{Natural Deduction}
+\begin{itemize}
+    \item \textbf{Natural Deduction} is a rule-based system for deriving conclusions from assumptions in logic.
+    \item Instead of using exhaustive truth tables, proofs are built step-by-step using inference rules.
+    \item Example: Proving from $A \land (A \rightarrow \bot)$ that $\bot$ (contradiction) can be derived.
+    \item Basis for how proof assistants check the logical structure of proofs.
+\end{itemize}
+\end{frame}
+\subsection*{Ins}
+\begin{frame}{Intuitionistic Logic}
+\begin{itemize}
+    \item \textbf{Intuitionistic Logic} formalizes constructive mathematics, where a statement is only true if a proof can be constructed.
+    \item Omits some classical logic principles, such as the Law of Excluded Middle.
+    \item Stronger requirement: to prove existence, a method or algorithm must be given.
+    \item Proof assistants leverage this constructive approach for digital verification.
+\end{itemize}
+\end{frame}
+\subsection*{$\lambda$-Calculus}
+\begin{frame}{$\lambda$-Calculus and Type Theory}
+\begin{itemize}
+    \item \textbf{$\lambda$-Calculus}: A foundational system for defining and applying functions using abstraction and application.
+    \item \textbf{Type Theory}: Assigns types to every term; ensures correctness of operations.
+    \item \textbf{Curry--Howard Correspondence}: 
+    \begin{itemize}
+        \item Propositions $\leftrightarrow$ Types
+        \item Proofs $\leftrightarrow$ Programs
+    \end{itemize}
+    \item \emph{Dependent types} allow types to depend on values, expressing complex logical properties.
+\end{itemize}
+\end{frame}
+\section{Architecture of proof assistant}
+\begin{frame}{Architecture of a Proof Assistant}
+\begin{itemize}
+    \item \textbf{Kernel}: Minimal, trustworthy codebase enforcing logical rules and validating proofs.
+    \item \textbf{Tactic Engine}: Helps build and automate proofs step by step.
+    \item \textbf{Formal Proof Language}: Rigorously expresses definitions, statements, and proofs.
+    \item \textbf{Libraries}: Collections of verified mathematical foundations for reuse.
+    \item \textbf{User Interface}: IDEs and plugins for interactive, efficient proof development.
+\end{itemize}
+\end{frame}
+\section{Comparative Study}
+\subsection*{Agda}
+\begin{frame}{Agda}
+\begin{itemize}
+    \item \textbf{Dependently typed functional programming language and proof assistant}.
+    \item No separate tactic layer---proofs are constructed explicitly and directly.
+    \item Uses \textbf{bidirectional type checking}, minimal kernel, and integrates normalization.
+    \item Emphasizes explicitness and formal clarity over automation.
+    \item Best suited for foundational work and educational purposes.
+\end{itemize}
+\end{frame}
+\subsection*{Rocq (Coq)}
+\begin{frame}{Rocq (Coq)}
+\begin{itemize}
+    \item \textbf{Tactic-based interactive theorem prover} famous for large formalizations (e.g., Four Color Theorem).
+    \item Kernel is based on Calculus of Inductive Constructions (CIC), written in Coq and extracted to OCaml.
+    \item Proofs often built with backward reasoning using tactics---automation is heavily supported.
+    \item Strong type-checking with bidirectional algorithms and conversion checking.
+    \item Widely used for complex, real-world formal verification tasks.
+\end{itemize}
+\end{frame}
+\subsection*{Lean}
+\begin{frame}{Lean}
+\begin{itemize}
+    \item Developed at Microsoft, supports both \textbf{tactic-based and term-based} proofs.
+    \item Kernel based on the Calculus of Inductive Constructions, implemented in C++/C.
+    \item Allows seamless switching between tactic and term styles---user-friendly and flexible.
+    \item Smart elaborator with coercion, backtracking, and overloading.
+    \item Popular in research and education, especially for combinatorial and mathematical formalizations.
+\end{itemize}
+\end{frame}
+\section{Formalization Of Some proofs}
+
 \input{AuxiliaryFiles/Thanks}  

### 2025-08-05 Mainpaper/paper.tex
REPORT DONE

--- 2025-08-05 Mainpaper/paper.tex
@@
 \DeclareUnicodeCharacter{03A3}{\ensuremath{\Sigma}}   % Σ
+\DeclareUnicodeCharacter{211A}{\ensuremath{\mathbb{Q}}} % ℚ
+\DeclareUnicodeCharacter{2223}{\ensuremath{\mid}}       % ∣
+\DeclareUnicodeCharacter{2264}{\ensuremath{\leq}}       % ≤
+\DeclareUnicodeCharacter{2261}{\ensuremath{\equiv}}     % ≡
+\DeclareUnicodeCharacter{21D2}{\ensuremath{\Rightarrow}}% ⇒
+\DeclareUnicodeCharacter{2243}{\ensuremath{\simeq}}     % ≃
+\DeclareUnicodeCharacter{2124}{\ensuremath{\mathbb{Z}}} % ℤ
+\DeclareUnicodeCharacter{2115}{\ensuremath{\mathbb{N}}} % ℕ
+\DeclareUnicodeCharacter{211D}{\ensuremath{\mathbb{R}}} % ℝ
+
 \geometry{
@@
         \includegraphics[width=4cm]{logo.png}\\[2cm]
-        % Replace with your logo file. Place 'ku_logo.png' in the same folder!
 
@@
 (Solane,1998) However the peer review took over 4 years especially due to the proof
-being incredibely difficult to check, with over a dozen of mathematician to refree
-the proof and although they were 99% certain it was not untill the proof became
+being incredibly difficult to check, with over a dozen of mathematician to referee 
+the proof and although they were 99 percent certain it was not until the proof became
 formalized that they were able to completely validate the correctness of the proof,
@@
 number of cases which is not feasible to write and verify manually and requires
-computer assitance.
+computer assistance.
 Such Proof assistants (systems) work by treating proofs as computational objects
@@
 ally, proof assistants ensure that every proof is exact, complete, and verifiable by
-a machine. This project invloves investigating this process.
+a machine. This project involves investigating this process.
 
@@
     \item To explore the correspondence between Proofs and Programs, Type Theory and Intuitionstic Logic.
-    \item To investigate a current exsisting proof assistants (Agda) and identify the mathematics behind it.
+    \item To investigate a current existing proof assistants (Agda) and identify the mathematics behind it.
     \item To formalize some selected proofs in Agda and demonstrate verification process.
@@
 
-\subsection{Defining Natural Numbers, Equality and Proof of Assocaitive and Commutative Properties}
+\subsection{Defining Natural Numbers, Equality and Proof of Associative and Commutative Properties}
 \begin{verbatim}
@@
 \end{verbatim}
-\subsection{Euclid's Theorem of Primes }
 
-\subsection{Euclid's Algorithm's Proof}
+\subsection{Attempt to Formalize Constructive Reals}
+\begin{verbatim}
+    module rls where
+open import Relation.Binary.Core using (Rel)
 
-\subsection{Attempt to Constructive Reals}
+open import Data.Rational as ℚ using (ℚ; -_ ; _*_; 
+  _-_; _+_; ∣_∣; _≤_; *≤* ; _≤?_; _÷_; ≡⇒≃)
+
+open import Data.Integer as ℤ using (ℤ; +_ ; -[1+_]; pos; _%_)
+open import Data.Nat as ℕ using (ℕ; suc; zero; _≤?_)
+import Level
+
+natToInteger : ℕ → ℤ
+natToInteger zero    = 0
+natToInteger (suc n) = pos (pos (suc n))
+
+oneℤ : ℤ
+oneℤ = pos 1
+
+reciprocal : ℕ → ℚ
+reciprocal zero    = 0 % 1
+reciprocal (suc n) = oneℤ % pos (suc n)
+
+record ℝ : Set where
+  constructor Real
+  field
+    f : ℕ -> ℚ
+    reg : {n m : ℕ} -> ∣ f n - f m ∣ ≤ reciprocal(suc n) + reciprocal(suc m)
+
+_≃_ : Rel ℝ Level.zero
+x ≃ y =  {n : ℕ} -> ∣ ℝ.f x n - ℝ.f y n ∣ ≤ reciprocal(suc n) + reciprocal(suc n)
+
+\end{verbatim}
 
 \section{Challenges and Workarounds}
+
+\begin{itemize}
+    \item In 5.1 we did everything from scratch, starting from definitons by ourselves.
+For larger and more proofs doing this for every .agda code is difficult.
+We make use of Agda Standard Library and leverage some pre defined objects and pre proven theorems.
+    \item In 5.3, For the second part of the therorem, we had take non constructive approach, which was done by using Dec from Relation.Nullary, which is the law of excluded middle.
+    \item In 5.4, We tried to construct real numbers as Cauchy sequences of Rational Numbers like in \cite{BB85}, where we encountered a problem while creating a reciprocal function $\mathbb{N} \to \mathbb{Q}$.
+The issue being in Agda the numerator requires to be a integer, and we did not find a way to introduce the integer 0.
+This remains unresolved.
+\end{itemize}
 

### 2025-08-05 Mainpaper/paper.tex
REPORT DONE1

@@
 
-\addcontentsline{toc}{section}{\textbf{ABSTRACT}}  
 \clearpage
@@
 \clearpage
+\addcontentsline{toc}{section}{\textbf{ABSTRACT}}
 
@@
 
+{\small
 \tableofcontents
+}
+\newpage
+\section*{Notations}
+
+Throughout this paper, we adopt notation commonly used in type theory and lambda calculus. For clarity, we summarize nonstandard or overloaded symbols below:
+
+\begin{itemize}
+
+  \item $\Gamma \vdash A$ denotes that the judgment $A$ is derivable from the context $\Gamma$. 
+  \item We occasionally write $A \subseteq B$ to mean $A$ implies $B$, in the sense of logical entailment. This should not be confused with the subset relation.
+  \item $A + B$ denotes the coproduct (disjoint union/ sum) of types $A$ and $B$, given A and B are types.
+  \item $\Sigma x : A.\ B(x)$ is the dependent sum type (also written $\sum$), generalizing pairs and existential quantification.
+  \item $\Pi x : A.\ B(x)$ denotes the dependent product type (also written $\prod$), generalizing function types and universal quantification.
+  \item $\lambda x.\ e$ denotes a function abstraction mapping $x$ to $e$.
+  \item $x \mapsto y$ is used informally to denote functions sending $x$ to $y$.
+  \item Inference rules and derivations are written using proof trees, with premises above a horizontal bar and conclusions below.
+\end{itemize}
 \newpage
@@
     \item To explore the correspondence between Proofs and Programs, Type Theory and Intuitionstic Logic.
-    \item To investigate a current existing proof assistants (Agda) and identify the mathematics behind it.
+    \item To investigate current existing proof assistants and identify the mathematics behind it.
     \item To formalize some selected proofs in Agda and demonstrate verification process.
@@
  The main focus is that this new set of rules does not contain the double negation rule which is present in what we introduced in Section 2.1.1.
-\textbf {Example \ref{lem}} uses the Double negation rule in its proof, which we don't have now.
+\textbf {Proposition \ref{lem}} uses the Double negation rule in its proof, which we don't have now.
 This agrees to that Law of Excluded Middle does not work in Intuitionstic Logic.
@@
 \end{enumerate}
-For comprehensive treatment of this topic, see \cite{Rojas2015}. And for the original formulation by Church \cite{}
+For comprehensive treatment of this topic, see \cite{Rojas2015}. And for the original formulation by Church \cite{church1936unsolvable}.
 
@@
 
-Every agda file has a top level module which corresponds to name of the file and the main programs goes inside the top level module.[]
-Another key component of agda is pattern matching. Pattern matching is a mechanism for analysing the structure of values. When inductively defined  datatypes are introduced the
-pattern matching becomes even more powerful as patten matching on one value can give the information about another value (thesis)
+Every agda file has a top level module which corresponds to name of the file and the main programs goes inside the top level module.
+Another key component of agda is pattern matching. Pattern matching is a mechanism for analysing the structure of values.
 
@@
 
-open import Data.Product using (Sigma; _,_)
+open import Data.Product using (Σ; _,_)
 
@@
 \end{itemize}
-
+\newpage
 % ======== References =========

### 2025-08-05 final-presentation/final.tex
presentations

@@
-\documentclass[aspectratio=169, compress, 12pt]{beamer}
+\documentclass[aspectratio=169, 12pt]{beamer}
 
@@
 \usepackage{AuxiliaryFiles/AuxiliaryFiles}  %%%%%%%%%%%%%    (DO NOT EDIT)
-
+\usepackage{graphics}
 \begin{document}
@@
 \end{frame}
-\section{Comparative Study}
-\subsection*{Agda}
-\begin{frame}{Agda}
+\subsection*{Kernel}
+\begin{frame}{Kernel: The Trusted Core}
 \begin{itemize}
-    \item \textbf{Dependently typed functional programming language and proof assistant}.
-    \item No separate tactic layer---proofs are constructed explicitly and directly.
-    \item Uses \textbf{bidirectional type checking}, minimal kernel, and integrates normalization.
-    \item Emphasizes explicitness and formal clarity over automation.
-    \item Best suited for foundational work and educational purposes.
+    \item The \textbf{kernel} is the minimal and most critical part of a proof assistant.
+    \item It enforces the logical rules of the underlying formal system (e.g., type theory).
+    \item Responsible for \textbf{validating every proof step} to guarantee correctness.
+    \item Ensures \textbf{soundness and trustworthiness}; the rest of the system depends on its integrity.
+    \item Typically very small and rigorously tested or formally verified to avoid bugs.
+    \item Example: Agda’s kernel is written in Haskell and integrates normalization to check definitional equality.
 \end{itemize}
 \end{frame}
-\subsection*{Rocq (Coq)}
-\begin{frame}{Rocq (Coq)}
+\subsection*{Tactic Engine}
+
+\begin{frame}{Tactic Engine: Proof Construction Assistant}
 \begin{itemize}
-    \item \textbf{Tactic-based interactive theorem prover} famous for large formalizations (e.g., Four Color Theorem).
-    \item Kernel is based on Calculus of Inductive Constructions (CIC), written in Coq and extracted to OCaml.
-    \item Proofs often built with backward reasoning using tactics---automation is heavily supported.
-    \item Strong type-checking with bidirectional algorithms and conversion checking.
-    \item Widely used for complex, real-world formal verification tasks.
+    \item The \textbf{tactic engine} supports users in constructing proofs interactively.
+    \item It breaks complex proof goals into simpler subgoals using \textbf{proof strategies} called tactics.
+    \item Provides \textbf{automation} for common proof patterns, speeding up proof development.
+    \item Enables both \textbf{forward} and \textbf{backward} reasoning approaches.
+    \item Even fully automated tactics rely on the kernel for final verification.
+    \item Varies among assistants (Agda has minimal/no tactics, Coq and Lean have powerful tactic systems).
 \end{itemize}
 \end{frame}
-\subsection*{Lean}
-\begin{frame}{Lean}
+\subsection*{ Language}
+
+\begin{frame}{Formal Proof Language: Expressing Proofs Precisely}
 \begin{itemize}
-    \item Developed at Microsoft, supports both \textbf{tactic-based and term-based} proofs.
-    \item Kernel based on the Calculus of Inductive Constructions, implemented in C++/C.
-    \item Allows seamless switching between tactic and term styles---user-friendly and flexible.
-    \item Smart elaborator with coercion, backtracking, and overloading.
-    \item Popular in research and education, especially for combinatorial and mathematical formalizations.
+    \item This language allows expressing \textbf{definitions, propositions, and proofs} rigorously.
+    \item Typically a \textbf{dependently typed language} so logical properties can be encoded as types.
+    \item Provides \textbf{syntax and semantics} suitable for formal reasoning and machine checking.
+    \item Enables users to write \textbf{human-readable yet unambiguous} formal proofs.
+    \item Integrates smoothly with tactics and type checker to maintain correctness.
+    \item Example languages: Agda’s core language, Coq’s Gallina, Lean’s dependent type language.
 \end{itemize}
 \end{frame}
+\subsection*{Libraries}
+\begin{frame}{Libraries: Reusable Verified Foundations}
+\begin{itemize}
+    \item Extensive collections of \textbf{formalized mathematics and algorithms} supporting new developments.
+    \item Include \textbf{basic theories} such as arithmetic, algebra, logic, and set theory.
+    \item Enable users to \textbf{build on existing verified results} without re-proving foundations.
+    \item Libraries evolve and grow, fostering \textbf{collaboration and community sharing}.
+    \item Well-maintained libraries reduce duplication and improve proof assistant adoption.
+    \item Examples include Coq’s Standard Library, Agda Standard Library, Lean’s mathlib.
+\end{itemize}
+\end{frame}
+\subsection*{User Interface}
+\begin{frame}{User Interface: Proof Development Environment}
+\begin{itemize}
+    \item Provides \textbf{interactive tools} like IDEs, editor plugins, or command line interfaces.
+    \item Features include \textbf{syntax highlighting, error reporting, real-time proof state visualization,} and \textbf{auto-completion}.
+    \item Enhances \textbf{usability and productivity} for proof authors.
+    \item Supports \textbf{integration with tactics and proof language} for seamless workflow.
+    \item Examples: CoqIDE, Proof General, Emacs-mode for Agda, VS Code extensions.
+    \item A good interface lowers the learning curve and makes formalization more accessible.
+\end{itemize}
+\end{frame}
+\section{Comparative Study}
+\begin{frame}{Comparative Table: Agda, Rocq (Coq), and Lean}
+
+\begin{table}
+\centering
+\scalebox{0.75}{
+    \begin{tabular}{|p{2cm}|p{4cm}|p{4cm}|p{4cm}|}
+    \hline
+    \small \textbf{Component} &  \small \textbf{Agda} & \small \textbf{Rocq (Coq)} & \small \textbf{Lean} \\
+    \toprule
+    \small \textbf{Proof Style} & \footnotesize Explicit term-based, manual proof writing & \footnotesize Tactic-based, automated backward reasoning & \footnotesize Both tactic-based and term-style \\
+
+    \small \textbf{Kernel} & Minimal, written in Haskell, tight integration with normalization & Based on Calculus of Inductive Constructions (CIC), written in Coq (extracted to OCaml) & CIC-based, written in C++/C \\
+    \small \textbf{Type Checking} & Bidirectional, transparent, normalization by evaluation & Bidirectional, heavy conversion, strong automation & Bidirectional, smart elaboration (coercion, backtracking, overloading) \\
+
+    \small \textbf{Automation} & Limited (no tactics, minimal automation) & Extensive tactic engine and proof search & Advanced, seamless tactic/term mixing, smart elaborator \\
+
+    \small \textbf{Use Cases} & Foundations, education, dependently typed programming & Large/complex formalizations, industrial-scale proofs & Research, education, combinatorial/mathematical formalizations \\
+    \bottomrule
+    \end{tabular}
+}
+\end{table}
+\end{frame}
+
 \section{Formalization Of Some proofs}
+\subsection*{}
 

### 2025-08-05 final-presentation/final.tex
pres

@@
 \begin{frame}{What are proof assistant}
-    proof assistant, are software more specifically a type of programming language thats allows us to 
+    Proof assistant, are software more specifically a type of programming language thats allows us to 
     formalize mathematical proofs in computer for digital verification. 
@@
     \begin{itemize}
-        \item faster computation for complex problems
-        \item many exceptional cases can be explored which would take mathematicians  long time \\
+        \item Fast and Efficient 
+        \item Many cases can be explored which would take mathematicians  long time \\
            \indent  \footnotesize ex:   The Kepler Conjecture's proof , which was so complex that verifying it manually would take 20 person-years, but proof assistants made this verification feasible and fast.
+
+        \item What if you don't use proof assistants? ABC conjecture
         \end{itemize}
+
 \end{frame}
-\section{Logical foundation}
+
 \begin{frame}
-    \begin{itemize}
-    \item Based on logic (natural deduction, intuitionistic logic), $\lambda$-calculus, and type theory.
-    \item Curry–Howard Correspondence:
-    \begin{itemize}
-        \item Propositions $\leftrightarrow$ Types
-        \item Proofs $\leftrightarrow$ Programs
-    \end{itemize}
-    \item Dependently Typed Languages: Types can depend on values, enabling encoding of properties and proofs.
-\end{itemize}
+    \begin{figure}[h]
+  \centering
+  \includegraphics[width=0.8\textwidth]{pa.jpeg}
+\end{figure}
 \end{frame}
+
+\section{Foundations}
+
 \subsection*{Naturl deduction}
@@
 \end{frame}
+\begin{frame}
+    \begin{figure}[h]
+  \centering
+  \includegraphics[width=0.8\textwidth]{nd.jpeg}
+\end{figure}
+\end{frame}
+
 \subsection*{Ins}
@@
 \begin{itemize}
-    \item \textbf{Intuitionistic Logic} formalizes constructive mathematics, where a statement is only true if a proof can be constructed.
-    \item Omits some classical logic principles, such as the Law of Excluded Middle.
+    \item \textbf{Intuitionistic Logic} Also called Constructive Logic, reflects principles of constructive mathematics, where a statement is only true if a proof can be constructed.
+    \item Omits some classical logic rules, such as the Law of Excluded Middle.
     \item Stronger requirement: to prove existence, a method or algorithm must be given.
@@
 \end{itemize}
+\end{frame}
+
+\begin{frame}
+    \begin{figure}[h]
+  \centering
+  \includegraphics[width=0.7\textwidth]{nid.jpeg}
+
+\end{figure}
+Inference Rules for Intuitionistic Logic
 \end{frame}
@@
     \item \textbf{Type Theory}: Assigns types to every term; ensures correctness of operations.
+         \item \emph{Dependent types} allow types to depend on values, expressing complex logical properties.
+
     \item \textbf{Curry--Howard Correspondence}: 
@@
     \end{itemize}
-    \item \emph{Dependent types} allow types to depend on values, expressing complex logical properties.
 \end{itemize}

### 2025-08-06 Mainpaper/paper.tex
presentations
minor edits

@@
 Proof assistants offer a dedicated programming language designed to express definitions, statements and proofs. These languages are typically Dependently typed,
-and allows implicit argument . they are highly readable and rigorous. Integration with tactic scripting enables a smooth blend of automated and manual proof development [Barras et al.,
-1997][Agda Team, 2023][Lean Development Group, 2024].
+and allows implicit argument . they are highly readable and rigorous. Integration with tactic scripting enables a smooth blend of automated and manual proof development \cite{}
 
@@
 Modern UIs include integrated development environments (IDEs) or editor plugins that provide syntax highlighting, error diagnostics, real-time proof state visualization,
-and automated tactic suggestion. Such interfaces significantly enhance productivity and make formal methods more accessible to a wider audience [CoqIDE, 2025][Isabelle/jEdit, 2023].
+and automated tactic suggestion. Such interfaces significantly enhance productivity and make formal methods more accessible to a wider audience \cite{isabelle2023}
 \subsection{Formal Libraries}
@@
 complex formalizations without reinventing foundational results.
-These libraries grow continuously and provide a shared repository of reusable components vital for large-scale verification projects [Coq Standard Library, 2025][Mathlib, 2025][Isabelle Archive of Formal Proofs].
+These libraries grow continuously and provide a shared repository of reusable components vital for large-scale verification projects \cite{rocqstdlib:v9.0.0}
 
@@
 In Dependently typed programming language there is no clear distinction between types and values , and types can depend on arbitary values and may appear as results of ordinary functions. 
-This feature of Dependently typed programs allows us to encode properties of values as type of proofs of that property, allowing us to use Dependent type programming as logic.
-(Dependently Typed Programming in Agda)
+This feature of Dependently typed programs allows us to encode properties of values as type of proofs of that property, allowing us to use Dependent type programming as logic. \cite{norell2008agda}
 
 Every agda file has a top level module which corresponds to name of the file and the main programs goes inside the top level module.
-Another key component of agda is pattern matching. Pattern matching is a mechanism for analysing the structure of values.
+Another key component of agda is pattern matching. Pattern matching is a mechanism for analysing the structure of values. \cite{norell2007thesis}
 
@@
 \end{itemize}
-\newpage
+\clearpage
+
 % ======== References =========

### 2025-08-06 final-presentation/final.tex
minor edits
presentation finalization

@@
 \usepackage{graphics}
+
+\DeclareUnicodeCharacter{2200}{\ensuremath{\forall}}    % ∀
+\DeclareUnicodeCharacter{2113}{\ensuremath{\ell}}       % ℓ
+\DeclareUnicodeCharacter{2192}{\ensuremath{\rightarrow}}% →
+\DeclareUnicodeCharacter{00D7}{\ensuremath{\times}}     % ×
+\DeclareUnicodeCharacter{228E}{\ensuremath{\uplus}}   % ⊎
+\DeclareUnicodeCharacter{22A5}{\ensuremath{\bot}}       % ⊥
+\DeclareUnicodeCharacter{00AC}{\ensuremath{\neg}}       % ¬
+\DeclareUnicodeCharacter{2081}{\textsubscript{1}}       % ₁
+\DeclareUnicodeCharacter{2082}{\textsubscript{2}}       % ₂
+\DeclareUnicodeCharacter{03BB}{\ensuremath{\lambda}}    % λ
+\DeclareUnicodeCharacter{03A3}{\ensuremath{\Sigma}}   % Σ
+\DeclareUnicodeCharacter{211A}{\ensuremath{\mathbb{Q}}} % ℚ
+\DeclareUnicodeCharacter{2223}{\ensuremath{\mid}}       % ∣
+\DeclareUnicodeCharacter{2264}{\ensuremath{\leq}}       % ≤
+\DeclareUnicodeCharacter{2261}{\ensuremath{\equiv}}     % ≡
+\DeclareUnicodeCharacter{21D2}{\ensuremath{\Rightarrow}}% ⇒
+\DeclareUnicodeCharacter{2243}{\ensuremath{\simeq}}     % ≃
+\DeclareUnicodeCharacter{2124}{\ensuremath{\mathbb{Z}}} % ℤ
+\DeclareUnicodeCharacter{2115}{\ensuremath{\mathbb{N}}} % ℕ
+\DeclareUnicodeCharacter{211D}{\ensuremath{\mathbb{R}}} % ℝ
 \begin{document}
@@
     \end{itemize}
+    \item \emph{Dependent types} allow types to depend on values, expressing complex logical properties.
 \end{itemize}
@@
 
-\section{Formalization Of Some proofs}
-\subsection*{}
+\section{Formalization Of Some Proofs}
+\subsection*{Defining Natural Numbers} 
+    \begin{frame}[fragile]
+        \frametitle{Eg: Defining Natural Numbers}
+        \begin{verbatim}
+           data N : Set where 
+                Zero : N 
+                suc : N -> N
+        \end{verbatim}
+    \end{frame}
+\subsection*{simple properities  } 
+\begin{frame}[fragile]
+    \frametitle{Eg: Some mathematical properities }
+    \begin{verbatim}
+        Transitivity properties:
+
+        data _==_ { A : Set } ( x : A ) : A -> Set where
+            refl : x == x
+
+        trans : { A : Set } { x y z : A}
+            -> x == y
+            -> y == z
+            -> x == z
+            trans refl refl = refl
+    \end{verbatim}
+\end{frame}
+\subsection{Formalization Of DeMorgan's Law} 
+\begin{frame}[fragile]
+    \frametitle{DeMorgan's Law in agda}
+   \footnotesize \begin{verbatim}
+DeMorgan’s Law
+open import Agda.Primitive using (Level; lzero)
+open import Data.Product using (_×_; _,_)
+open import Data.Sum using (_⊎_; inj1; inj2)
+open import Relation.Nullary using (Dec; yes; no)
+open import Data.Empty using (⊥; ⊥-elim)
+open import Relation.Nullary.Negation using (¬_)
+--One Direction
+deMorganOneWay : ∀ {ℓ} {P Q : Set ℓ} → (¬ P) ⊎ (¬ Q) → ¬ (P × Q)
+deMorganOneWay (inj1 np) (p , q) = np p
+deMorganOneWay (inj2 nq) (p , q) = nq q
+
+    \end{verbatim}
+
+\end{frame}
+\begin{frame}[fragile]
+
+    \small \begin{verbatim}
+Converse, Requires Non Constructive Assumptions
+deMorganOtherWay :
+∀ {ℓ} {P Q : Set ℓ}
+→ Dec P
+→ Dec Q
+→ ¬ (P × Q)
+→ (¬ P) ⊎ (¬ Q)
+deMorganOtherWay (yes p) _ notPQ = inj2 (λ q → notPQ (p , q))
+deMorganOtherWay (no np) (yes q) _ = inj1 np
+deMorganOtherWay (no np) (no nq) _ = inj1 np -- or inj2 nq
+    \end{verbatim}
+
+\end{frame}
+\section{Limitations}
+\begin{frame}{Limitations}
+    \begin{itemize}
+        \item Unable to completely formalize " Constructive Reals" 
+        \item issue encountered while creating a reciprocal function N → Q  
+        \item Limited time to explore the rigorous algorithm of bi-directional type checking
+
+    \end{itemize}
+\end{frame}
 

### 2025-08-06 final-presentation/final.tex

--- 2025-08-05 final-presentation/final.tex
@@
 
-\subsection*{Naturl deduction}
 \begin{frame}{Natural Deduction}
@@
     \item Example: Proving from $A \land (A \rightarrow \bot)$ that $\bot$ (contradiction) can be derived.
-    \item Basis for how proof assistants check the logical structure of proofs.
 \end{itemize}
@@
 
-\subsection*{Ins}
 \begin{frame}{Intuitionistic Logic}
@@
 \end{frame}
-\subsection*{$\lambda$-Calculus}
-\begin{frame}{$\lambda$-Calculus and Type Theory}
+\begin{frame}{Lambda Calculus}
+\textbf{What is Lambda Calculus?}
 \begin{itemize}
-    \item \textbf{$\lambda$-Calculus}: A foundational system for defining and applying functions using abstraction and application.
-    \item \textbf{Type Theory}: Assigns types to every term; ensures correctness of operations.
-         \item \emph{Dependent types} allow types to depend on values, expressing complex logical properties.
-
-    \item \textbf{Curry--Howard Correspondence}: 
+  \item A formal model of computation by Alonzo Church
+  \item Lambda Terms:
     \begin{itemize}
-        \item Propositions $\leftrightarrow$ Types
-        \item Proofs $\leftrightarrow$ Programs
+        \item Variables: $x$, $y$, $z$
+        \item Abstraction: $\lambda x.E$
+        \item Application: $(\lambda x.E)F$
     \end{itemize}
@@
 \end{frame}
+\begin{frame}
+\textbf{Examples}
+\begin{itemize}
+    \item $\lambda x.x^2$ is a function
+    \item $(\lambda x.x^2)(3) \rightarrow 9$
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Type Theory \& Simply Typed Lambda Calculus}
+\textbf{Type Theory Basics}
+\begin{itemize}
+  \item Assigns types to terms: $1 : \mathbb{N}$, $+ : \mathbb{N} \times \mathbb{N} \to \mathbb{N}$
+  \item Typing is decidable
+\end{itemize}
+
+\textbf{Type Categories}
+\begin{itemize}
+  \item Base Types: $\mathbb{N}$, Bool, $\bot$
+  \item Arrow Types: $f : A \to B$
+  \item Product Types: $\langle a, b \rangle : A \times B$
+  \item Sum Types: $a : A + B$
+\end{itemize}
+\end{frame}
+\begin{frame}
+   \textbf{Typed Lambda Calculus}
+\begin{itemize}
+  \item $(\lambda x : \mathbb{N}.x^2) : \mathbb{N} \to \mathbb{N}$
+\end{itemize}
+ \textbf{Dependent Types}
+\begin{itemize}
+  \item Types depend on values: $Vec(n)$
+  \item Indexed types, predicate representation
+  \item $Vec : \mathbb{N} \to Type$
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{ Curry–Howard Correspondence}
+    \begin{itemize}
+  \item sometimes referred as Curry Howard Isomorphism
+  \item A connection between logic, computation, and type theory
+  \item Also known as the \textit{proofs-as-programs} and \textit{propositions-as-types} principle
+\end{itemize}
+
+\textbf{Core Idea}
+\begin{itemize}
+  \item A \textbf{proposition} corresponds to a \textbf{type}
+  \item A \textbf{proof} of the proposition is a \textbf{program} (term) of that type
+  \item Proof checking is equivalent to type checking
+  \item Proof normalization corresponds to program evaluation
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\textbf{Curry–Howard Correspondence}
+\begin{tabular}{|c|c|c|}
+\hline
+Logic & Type Theory & Programming \\
+\hline
+Proposition & Type & - \\
+Proof & Term & Program \\
+Implication $A \to B$ & $A \to B$ & $\lambda$-abstraction \\
+Conjunction $A \land B$ & $A \times B$ & Pair \\
+Disjunction $A \lor B$ & $A + B$ & Tagged union \\
+Falsehood $\bot$ & Empty Type & No term \\
+Universal $\forall x.A(x)$ & $\Pi x:A.B(x)$ & Function over types \\
+Existential $\exists x.A(x)$ & $\Sigma x:A.B(x)$ & Dependent pair \\
+\hline
+\end{tabular}
+
+\end{frame}
+
+    \begin{frame}{Martin-Löf Type Theory}
+\textbf{What is it?}
+\begin{itemize}
+  \item A formal system for constructive mathematics
+  \item Also known as Intuitionistic Type Theory
+  \item Backbone of modern Proof Assistants
+\end{itemize}
+
+\textbf{Core Types}
+\begin{itemize}
+  \item \textbf{} ($\Pi$-type): Dependent function type, $\forall x:A.B(x)$
+  \item \textbf{} ($\Sigma$-type): Dependent sum type, $\exists x:A.B(x)$
+  \item \textbf{Identity Type}: Internal equality between terms
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+  \textbf{Type Universes}
+\begin{itemize}
+  \item Types have types: $1 : \mathbb{N}, \mathbb{N} : Type, Type : \cdots$
+  \item Not like sets within sets — avoids paradoxes
+  \item Enables reasoning and abstraction over types themselves
+
+\end{itemize}  
+\end{frame}
+
 \section{Architecture of proof assistant}
@@
 \section{Formalization Of Some proofs}
-\subsection*{}
+\subsection*{Defining Natural Numbers} 
+    \begin{frame}[fragile]
+        \frametitle{Eg: Defining Natural Numbers}
+        \begin{verbatim}
+           data N : Set where 
+                Zero : N 
+                suc : N -> N
+        \end{verbatim}
+    \end{frame}
 

### 2025-08-06 final-presentation/final.tex

@@
 
-\section{Formalization Of Some proofs}
+\section{Formalization Of Some Proofs}
 \subsection*{Defining Natural Numbers} 
@@
     \end{frame}
+\subsection*{simple properities  } 
+\begin{frame}[fragile]
+    \frametitle{Eg: Some mathematical properities }
+    \begin{verbatim}
+        Equality and Transitivity:
+
+        data _==_ { A : Set } ( x : A ) : A -> Set where
+            refl : x == x
+
+        trans : { A : Set } { x y z : A}
+            -> x == y
+            -> y == z
+            -> x == z
+            trans refl refl = refl
+    \end{verbatim}
+\end{frame}
+\subsection{Formalization Of DeMorgan's Law} 
+\begin{frame}[fragile]
+    \frametitle{DeMorgan's Law in agda}
+   \footnotesize \begin{verbatim}
+DeMorgan’s Law
+open import Agda.Primitive using (Level; lzero)
+open import Data.Product using (_×_; _,_)
+open import Data.Sum using (_⊎_; inj1; inj2)
+open import Relation.Nullary using (Dec; yes; no)
+open import Data.Empty using (⊥; ⊥-elim)
+open import Relation.Nullary.Negation using (¬_)
+--One Direction
+deMorganOneWay : ∀ {ℓ} {P Q : Set ℓ} → (¬ P) ⊎ (¬ Q) → ¬ (P × Q)
+deMorganOneWay (inj1 np) (p , q) = np p
+deMorganOneWay (inj2 nq) (p , q) = nq q
+
+    \end{verbatim}
+
+\end{frame}
+\begin{frame}[fragile]
+
+    \small \begin{verbatim}
+Converse, Requires Non Constructive Assumptions
+deMorganOtherWay :
+∀ {ℓ} {P Q : Set ℓ}
+→ Dec P
+→ Dec Q
+→ ¬ (P × Q)
+→ (¬ P) ⊎ (¬ Q)
+deMorganOtherWay (yes p) _ notPQ = inj2 (λ q → notPQ (p , q))
+deMorganOtherWay (no np) (yes q) _ = inj1 np
+deMorganOtherWay (no np) (no nq) _ = inj1 np -- or inj2 nq
+    \end{verbatim}
+
+\end{frame}
+\section{ Limitations}
+\begin{frame}{}
+    \begin{itemize}
+        \item Issue while implementing Real Numbers Construcitvely.
+        \item Bidirectional Typechecking Algorithm isn't discussed rigorously.
+
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{References}
+\footnotesize
+\begin{itemize}
+  \item William Howard. \textit{The formula-as-types notion of construction}, 1969.
+  \item Per Martin-Löf. \textit{An Intuitionistic Theory of Types}, 1972.
+  \item Jan Willem Klop, Alejandro Ríos. \textit{Introduction to Lambda Calculus}, Rojas, 2015.
+  \item Herman Geuvers. \textit{Proof Assistants: History, Ideas and Future}, 2009.
+  \item Frank Pfenning. \textit{Intuitionistic Logic: Proof Theory}, Lecture Notes, 2004.
+ \item Sozeau, M. et al. (2025). \textit{Correct and complete type checking and certified erasure for Coq, in Coq}, J. ACM, 72(1).
+  \item Team, R. P. D. (2025). \textit{Rocq Prover Reference Manual: Core Language}. Accessed: 2025-08-05.
+\end{itemize}
+\end{frame}
 

### 2025-08-06 Mainpaper/paper.tex
spells?
intspeelling

@@
 \begin{enumerate}
-    \item To explore the correspondence between Proofs and Programs, Type Theory and Intuitionstic Logic.
+    \item To explore the correspondence between Proofs and Programs, Type Theory and Intuitionistic Logic.
     \item To investigate current existing proof assistants and identify the mathematics behind it.
@@
 The \textbf{soundness} and \textbf{completeness} of this system are discussed in Section 3.1 and 3.2 \cite{Alrubyli2021}. 
-\subsubsection{Intuitionstic Logic}
-Intuitionstic Logic was introduced to formalize the constructive method to do mathematics.
+\subsubsection{Intuitionistic Logic}
+Intuitionistic Logic was introduced to formalize the constructive method to do mathematics.
 Unlike in Classical Logic a statement is True if we can construct a proof for it and to claim a statement is False, again a proof of its falsity is required.
@@
 To show something exists one must provide an method or algorithm to construct it. Proof Assistants leverage this fact.
-The constructive view of doing mathematics gives a stricter criteria. Intuitionstic  Logic can be obtained by restricting certain parts of Classical Logic, like the Law of Excluded Middle.
+The constructive view of doing mathematics gives a stricter criteria. Intuitionistic  Logic can be obtained by restricting certain parts of Classical Logic, like the Law of Excluded Middle.
 
-For inference rules for Natural Deduction in Intuitionstic Logic See 2.1 \cite{Pfenning2004}, here the language is slightly different from above, we have
+For inference rules for Natural Deduction in Intuitionistic Logic See 2.1 \cite{Pfenning2004}, here the language is slightly different from above, we have
 \\
@@
 \textbf {Proposition \ref{lem}} uses the Double negation rule in its proof, which we don't have now.
-This agrees to that Law of Excluded Middle does not work in Intuitionstic Logic.
+This agrees to that Law of Excluded Middle does not work in Intuitionistic Logic.
 The method of proof by contradiction also relies on this rule, so constructivists omit it.\\
@@
     Let us treat every proposition or Logical Formula as a Type, and their proofs as the term inhabiting it.
-Intuitionstically, we interpret a proposition being true if we present a proof for it.
+Intuitionistically, we interpret a proposition being true if we present a proof for it.
 Here, it will be true if there is a term (proof) inhabiting the type (proposition).
-    p:P is now interpreted as p is a proof of P. In \cite{How} the strong correspondence between intuitionstic derivation and lambda terms was mentioned. In essence the corrrespondence is as follows,
+    p:P is now interpreted as p is a proof of P. In \cite{How} the strong correspondence between intuitionistic derivation and lambda terms was mentioned. In essence the corrrespondence is as follows,
     \begin{table}[h]
@@
 The use of dependent product and sum for the Quantifier highlights the necessity of Dependent Type Theory.
-The rules  (Proof Terms only), equivalent to Deduction rules of Intuitionstic Logic are in Section 2.4 of \cite{Pfenning2004}.
+The rules  (Proof Terms only), equivalent to Deduction rules of Intuitionistic Logic are in Section 2.4 of \cite{Pfenning2004}.
 Note that this does not contain the rule for predicates.
@@
     Martin-Löf Type Theory is the formal logic system that was developed to support constructive mathematics.
-It's often referred to as Intuitionstic Type Theory.
+It's often referred to as Intuitionistic Type Theory.
 It carries the use of dependent types and is the backbone of Proof Assistants.\\
